---
title: "CRBM Variables"
output: html_document
date: "2024-10-28"
---

# Cognitive Reserve and Brain Maintenance Variables

This document provides instructions and code for calculating cognitive reserve and brain maintenance variables based on the annual changes in a composite episodic memory variable and hippocampal volume. While this document is focused on hippocampal volume, the code can be adapted to incorporate other measures (e.g., cingulate cortex thickness or a composite measure). It is important that any new measure captures longitudinal changes in brain morphology to ensure it accurately represents brain maintenance.

## Recreating Cognitive Reserve and Brain Maintenance Variables

To recreate these variables using episodic memory and hippocampal volume, follow the provided code below. If you are using different measures, please modify the code accordingly to create an updated version of the cognitive reserve and brain maintenance variables.

### Data Requirements

- At least two time points of data
- Episodic memory variable/s
- Brain structure variable (e.g., hippocampal volume)
- Participant ages

```{r setup, include=FALSE}
if (!require("ggplot2")) {
  install.packages("ggplot2")
  require("ggplot2")
}
if (!require("dplyr")) {
  install.packages("dplyr")
  require("dplyr")
}
if (!require("patchwork")) {
  install.packages("patchwork")
  require("patchwork")
}
if (!require("readr")) {
  install.packages("readr")
  require("readr")
}
```

### Step 1: Load Your Data

Begin by importing your dataset. Ensure the data contains either:

1. Episodic memory variables, hippocampal volume, timepoints, and age
2. Previously calculated episodic memory composite (scaled by baseline) and episodic memory slopes plus hippocampal slopes

It is best to read in your data as long format if using the first variables and in wide format if using the second variables.

```{r warning=FALSE}
data <- read.csv("your data.csv")
```

### Step 2a: Calculate Slopes (if needed)

If your data are in long format and you have not yet calculated a composite episodic memory variable or each variable's slopes, use the following code `{r calculate slopes}` to generate these. Customize your variable names and specify your number of timepoints accordingly. 

```{r calculate slopes}
# Calculate Slopes: Insert variable names and timepoint number

# Define your memory variables
memory_vars <- c("memory variable 1", "memory variable 2") # Add more as needed

# Specify the number of timepoints
n_timepoints <- 6

# Rename your variables directly
data <- data %>%
  rename(
    hippocampus_volume = "your hippocampus volume variable",
    timepoint = "your timepoint variable", # This variable needs to be numeric (e.g. timepoint = 1 not timepoint = R01)
    id = "your id variable",
    age = "your age variable"
  )

# Optionally, uncomment to filter to only include the needed variables for this analysis
# data <- data %>% select(id, age, timepoint, hippocampus_volume, starts_with("memory_var_")) # Match `memory_vars`

# Main processing function
process_data <- function(data, n_timepoints, mem_vars) {
  # Convert data to wide format
  data_wide <- data
  for (i in 1:n_timepoints) {
    suffix <- paste0("_", i)
    timepoint_data <- data %>%
      filter(timepoint == i) %>%
      rename_with(~ paste0(.x, suffix), -c(id))

    if (i == 1) {
      data_wide <- timepoint_data
    } else {
      data_wide <- full_join(data_wide, timepoint_data, by = "id")
    }
  }

  # Remove timepoint column
  data_wide <- data_wide %>% select(-starts_with("timepoint_"))

  # Function to find the first non-NA value
  get_first_non_na <- function(x) {
    non_na_values <- x[!is.na(x)]
    if (length(non_na_values) > 0) {
      return(non_na_values[1])
    } else {
      return(NA)
    }
  }

  # Generate baseline variables
  for (j in seq_along(mem_vars)) {
    mem_var <- mem_vars[j]
    baseline_name <- paste0(mem_var, "_baseline")
    data_wide[[baseline_name]] <- apply(
      data_wide[, paste0(mem_var, "_", 1:n_timepoints), drop = FALSE],
      1,
      get_first_non_na
    )
  }

  # Scale memory variables by baseline (timepoint 1)
  for (i in 1:n_timepoints) {
    for (j in seq_along(mem_vars)) {
      mem_var <- mem_vars[j]
      base_col_name <- paste0(mem_var, "_baseline")

      # Current timepoint memory variable column
      col_name <- paste0(mem_var, "_", i)

      # Check if column exists
      if (col_name %in% names(data_wide)) {
        # Calculate baseline mean and SD
        baseline_mean <- mean(data_wide[[base_col_name]], na.rm = TRUE)
        baseline_sd <- sd(data_wide[[base_col_name]], na.rm = TRUE)

        # Scale the variable using baseline mean and SD
        scaled_values <- (data_wide[[col_name]] - baseline_mean) / baseline_sd
        data_wide[[paste0("sc_", col_name)]] <- as.numeric(scaled_values) # Convert to numeric
      } else {
        cat("Column not found:", col_name, "\n")
      }
    }
  }

  # Calculate memory composites
  for (i in 1:n_timepoints) {
    mem_cols <- paste0("sc_", mem_vars, "_", i)
    data_wide <- data_wide %>%
      mutate(!!paste0("memory_", i) := rowMeans(
        select(., any_of(mem_cols)),
        na.rm = TRUE
      ))
  }

  # Calculate slopes
  for (i in 1:nrow(data_wide)) {
    em_data <- data_wide[i, paste0("memory_", 1:n_timepoints)]
    hc_data <- data_wide[i, paste0("hippocampus_volume_", 1:n_timepoints)]
    age <- data_wide[i, paste0("age_", 1:n_timepoints)]

    if (sum(!is.na(em_data)) > 1 && sum(!is.na(hc_data)) > 1) {
      data_wide[i, "memory_slopes"] <- lm(em_data[!is.na(em_data)] ~ age[!is.na(em_data)])$coeff[2]
      data_wide[i, "hippocampus_slopes"] <- lm(hc_data[!is.na(hc_data)] ~ age[!is.na(hc_data)])$coeff[2]
    }
  }

  data_wide
}

# Execute the process on your data
data <- process_data(data, n_timepoints, memory_vars)
```

### Step 2b: Define Relevant Variables (if needed)

For those with pre-calculated hippocampus and memory slopes, make sure your data aligns with the naming conventions in this script. Your data should already be in wide format.

```{r define variables}
# Define your variables
data <- data %>%
  rename(
    hippocampus_slopes = "hippocampus slopes variable",
    memory_slopes = "memory slopes variable"
  )
```

### Step 3: Further Processing

With your standard variable names and required calculations complete, you can run the rest of this code without needed to make any more modifications. 

This block of code calculated the residuals of a regression between memory change and baseline age and another between hippocampal change and baseline age to be used in the calculation of the brain maintenance and cognitive reserve variables.

```{r}
# Remove NAs 
data <- data %>%
  drop_na(hippocampus_slopes, memory_slopes)

# Fit linear models with age at baseline as a covariate
mod_memory <- lm(memory_slopes ~ age_1, data = data)
mod_hippocampus <- lm(hippocampus_slopes ~ age_1, data = data)

# Attach residuals to the data frame
data <- data %>%
  mutate(
    residuals_memory_age = resid(mod_memory),
    residuals_hippocampus_age = resid(mod_hippocampus)
  )
```
```

This creates a scaled version of the hippocampus and memory annual change variables.

```{r}
data <- data %>%
  mutate(
    hippocampus_slopes_scaled = residuals_hippocampus_age / max(
      abs(min(residuals_hippocampus_age, na.rm = TRUE)),
      abs(max(residuals_hippocampus_age, na.rm = TRUE))
    ),
    memory_slopes_scaled = residuals_memory_age / max(
      abs(min(residuals_memory_age, na.rm = TRUE)),
      abs(max(residuals_memory_age, na.rm = TRUE))
    )
  )
```

Then this creates the brain maintenance (BM) variable.

```{r}
# Define a function to calculate the distance from a given data point to the point (1,1)
distance_to_bm_point <- function(y, x) {
  # Calculate the shortest distance from the data point to (1,1)
  distance <- sqrt((x - 1)^2 + (y - 1)^2)

  return(distance)
}

# Define a function to calculate the distance from a given data point to the point (1,1)
distance_to_bm_line <- function(y, x) {
  # Calculate the intersection point (projection of the point onto the line x = y)
  intersection_x <- (x + y) / 2
  intersection_y <- intersection_x

  # Calculate the distance from the point (x, y) to the projection point
  distance <- sqrt((intersection_x - x)^2 + (intersection_y - y)^2)

  return(distance)
}

# Calculate BM variable
data <- data %>%
  mutate(
    dist_to_bm_point = distance_to_bm_point(memory_slopes_scaled, hippocampus_slopes_scaled),
    dist_to_bm_line = distance_to_bm_line(memory_slopes_scaled, hippocampus_slopes_scaled),
    brain_maintenance = 3.414214 - (dist_to_bm_point + dist_to_bm_line)
  )
```

Next, this creates the cognitive reserve (CR) variable.

```{r}
# Define a function to calculate the distance from the data point to the point (1,1)
distance_to_cr_point <- function(y, x) {
  # Calculate the shortest distance from the data point to (1,1)
  distance <- sqrt((x + 1)^2 + (y - 1)^2)

  return(distance)
}

# Define a function to calculate the distance from to the line x = -y
distance_to_cr_line <- function(y, x) {
  # Calculate the intersection point (projection of the point onto the line x = -y)
  intersection_x <- (x - y) / 2
  intersection_y <- -(intersection_x)

  # Calculate the distance from the point (x, y) to the intersection point
  distance <- sqrt((intersection_x - x)^2 + (intersection_y - y)^2)

  return(distance)
}

# Calculate CR variable
data <- data %>%
  mutate(
    dist_to_cr_point = distance_to_cr_point(memory_slopes_scaled, hippocampus_slopes_scaled),
    dist_to_cr_line = distance_to_cr_line(memory_slopes_scaled, hippocampus_slopes_scaled),
    cognitive_reserve = 3.414214 - (dist_to_cr_point + dist_to_cr_line)
  )
```

Lastly, this calculates whether someone is aging more brain maintenance or cognitive reserve-like.

```{r}
# Classify as on the BM or CR aging pathway
data <- data %>%
  mutate(
    pathway = ifelse(dist_to_cr_line > dist_to_bm_line, "Brain Maintenance Pathway", "Cognitive Reserve Pathway")
  )
```


