---
title: "Analysis"
output: html_document
date: "2024-04-30"
---
```{r setup, include=FALSE}
if (!require("ggplot2")) {
  install.packages("ggplot2")
  require("ggplot2")
}
if (!require("dplyr")) {
  install.packages("dplyr")
  require("dplyr")
}
if (!require("psych")) {
  install.packages("psych")
  require("psych")
}
if (!require("tidyr")) {
  install.packages("tidyr")
  require("tidyr")
}
```

Start by reading in the clean data in wide format 
```{r warning=FALSE}
clean_data <- read.csv("/tsd/p274/data/durable/projects/p027-cr_bm/Clean Data All Cohorts/Wide Format/Merged cohorts analysed data wide.csv")
```

Visualize the HC and memory data
```{r warning=FALSE}
# Look at memory slopes
ggplot(clean_data, aes(x = memory_slopes)) +
  geom_histogram()

memory_normality <- shapiro.test(clean_data$memory_slopes)
print(memory_normality)

# Look at hippocampus slopes
ggplot(clean_data, aes(x = hc_slopes)) +
  geom_histogram()

hc_normality <- shapiro.test(clean_data$hc_slopes)
print(hc_normality)

# Visualize the FC data
ggplot(clean_data, aes(x = dmn_slopes)) +
  geom_histogram()

dmn_normality <- shapiro.test(clean_data$dmn_slopes)
print(dmn_normality)
```
Fit residual models to be able to use in the figures throughout the results section
```{r}
# In order to be able to save these models, the NAs have to be removed so I made a df for each type to remove the data and then can merge back into clean_df

# Create a memory-residual df
mem_res_data <- clean_data %>%
  drop_na(memory_slopes) %>%
  mutate(
    res_mem_slopes_age = residuals(
      lm(memory_slopes ~ age_followup, data = .)
    )
  ) %>%
  select(id, res_mem_slopes_age)

# Create a hippocampus-residual df
hc_res_data <- clean_data %>%
  drop_na(hc_slopes) %>%
  mutate(
    res_hc_slopes_age = residuals(
      lm(hc_slopes ~ age_followup, data = .)
    )
  ) %>%
  select(id, res_hc_slopes_age)

# Merge the two residual columns back together by id
emhc_data <- merge(mem_res_data, hc_res_data, by = c("id"), all = TRUE)
emhc_data <- merge(clean_data, emhc_data, by = c("id"), all = TRUE)
```

Using the new formula to calculate a BM continuous variable 
```{r}
# First create a scaled version of the HC and memory annual change variables using the residuals
emhc_data <- emhc_data %>%
  mutate(
    res_hc_slopes_scaled =
      res_hc_slopes_age / max(
        abs(min(res_hc_slopes_age, na.rm = TRUE)),
        abs(max(res_hc_slopes_age, na.rm = TRUE))
      ),
    res_mem_slopes_scaled =
      res_mem_slopes_age / max(
        abs(min(res_mem_slopes_age, na.rm = TRUE)),
        abs(max(res_mem_slopes_age, na.rm = TRUE))
      )
  )

# Define the function to calculate the distance from the data point to the point (1,1)
distance_to_bm_point <- function(y, x) {
  # Calculate the shortest distance from the data point to (1,1)
  distance <- sqrt((x - 1)^2 + (y - 1)^2)

  return(distance)
}

emhc_data <- emhc_data %>%
  mutate(dist_to_bm_point = distance_to_bm_point(res_mem_slopes_scaled, res_hc_slopes_scaled))

# Define the function to calculate the distance from the data point to the point (1,1)
distance_to_bm_line <- function(y, x) {
  # Calculate the intersection point (projection of the point onto the line x = y)
  intersection_x <- (x + y) / 2
  intersection_y <- intersection_x

  # Calculate the distance from the point (x, y) to the projection point
  distance <- sqrt((intersection_x - x)^2 + (intersection_y - y)^2)

  return(distance)
}

emhc_data <- emhc_data %>%
  mutate(dist_to_bm_line = distance_to_bm_line(res_mem_slopes_scaled, res_hc_slopes_scaled))

emhc_data <- emhc_data %>%
  mutate(dist_to_double_bm = dist_to_bm_line + dist_to_bm_point)
```

Descriptive stats BM
```{r}
describe(emhc_data$dist_to_double_bm)
shapiro.test(emhc_data$dist_to_double_bm)
ggplot(emhc_data, aes(x = dist_to_double_bm)) +
  geom_histogram(fill = "lightpink", color = "black", binwidth = 0.1)
```

Now create the CR continuous variable 
```{r}
# Define the function to calculate the distance from the data point to the point (1,1)
distance_to_cr_point <- function(y, x) {
  # Calculate the shortest distance from the data point to (1,1)
  distance <- sqrt((x + 1)^2 + (y - 1)^2)

  return(distance)
}


emhc_data <- emhc_data %>%
  mutate(dist_to_cr_point = distance_to_cr_point(res_mem_slopes_scaled, res_hc_slopes_scaled))

# Define the function to calculate the distance from to the line x = -y
distance_to_cr_line <- function(y, x) {
  # Calculate the intersection point (projection of the point onto the line x = -y)
  intersection_x <- (x - y) / 2
  intersection_y <- -(intersection_x)

  # Calculate the distance from the point (x, y) to the intersection point
  distance <- sqrt((intersection_x - x)^2 + (intersection_y - y)^2)

  return(distance)
}

emhc_data <- emhc_data %>%
  mutate(dist_to_cr_line = distance_to_cr_line(res_mem_slopes_scaled, res_hc_slopes_scaled))


emhc_data <- emhc_data %>%
  mutate(dist_to_double_cr = dist_to_cr_line + dist_to_cr_point)
```

Descriptive stats CR
```{r}
describe(emhc_data$dist_to_double_cr)
shapiro.test(emhc_data$dist_to_double_cr)
ggplot(emhc_data, aes(x = dist_to_double_cr)) +
  geom_histogram(fill = "lightpink", color = "black", binwidth = 0.1)
```

Now inverse BM and CR so that high values are good
```{r}
# Define function to reverse the CR / BM variables using their possible min and max values
reverse_values <- function(values) {
  max_value <- 3.414214
  min_value <- 0
  reversed <- (max_value + min_value) - values
  return(reversed)
}

# Reverse the bm and cr values
emhc_data$reverse_bm <- reverse_values(emhc_data$dist_to_double_bm)
emhc_data$reverse_cr <- reverse_values(emhc_data$dist_to_double_cr)

# Check that the output is ok
reversed <- emhc_data %>%
  select(id, reverse_bm, dist_to_double_bm, reverse_cr, dist_to_double_cr)
```

Now write the data frames as CSVs
```{r}
write.csv(emhc_data, file = "/tsd/p274/data/durable/projects/p027-cr_bm/Clean Data All Cohorts/Wide Format/Analysed data.csv")
```
