---
title: "Results"
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
if (!require("ggplot2")) {
  install.packages("ggplot2")
  require("ggplot2")
}
if (!require("dplyr")) {
  install.packages("dplyr")
  require("dplyr")
}
if (!require("psych")) {
  install.packages("psych")
  require("psych")
}
if (!require("stringr")) {
  install.packages("stringr")
  require("stringr")
}
if (!require("tidyr")) {
  install.packages("tidyr")
  require("tidyr")
}
if (!require("kableExtra")) {
  install.packages("kableExtra")
  require("kableExtra")
}
if (!require("knitr")) {
  install.packages("knitr")
  require("knitr")
}
if (!require("scales")) {
  install.packages("scales")
  require("scales")
}
if (!require("patchwork")) {
  install.packages("patchwork")
  require("patchwork")
}
if (!require("mgcv")) {
  install.packages("mgcv")
  require("mgcv")
}
if (!require("readr")) {
  install.packages("readr")
  require("readr")
}
if (!require("readxl")) {
  install.packages("readxl")
  require("readxl")
}
if (!require("lme4")) {
  install.packages("lme4")
  require("lme4")
}
if (!require("lmerTest")) {
  install.packages("lmerTest")
  require("lmerTest")
}
if (!require("gamm4")) {
  install.packages("gamm4")
  require("gamm4")
}
```

```{r echo=FALSE, warning=FALSE}
clean_data <- read.csv("/tsd/p274/data/durable/projects/p027-cr_bm/Clean Data All Cohorts/Wide Format/Analysed data.csv")

p_value_df <- read.csv("/tsd/p274/data/durable/projects/p027-cr_bm/p_value.csv")

# Change the p-values to none scientific format
p_value_df$adjusted_p_value <- as.numeric(format(p_value_df$adjusted_p_value, scientific = FALSE))
```

```{r echo=FALSE}
participants <- clean_data %>% nrow()
female <- (clean_data %>% filter(sex == "female") %>% nrow())
female_percentage <- (female / participants) * 100
mean_age <- mean(clean_data$age_baseline, na.rm = TRUE)
sd_age <- sd(clean_data$age_baseline, na.rm = TRUE)

# Create a new variable to represent average follow-up time
clean_data <- clean_data %>%
  mutate(FU_time = rowMeans(select(., mem_time, hc_time, fc_time), na.rm = TRUE))

mean_fu <- mean(clean_data$FU_time, na.rm = TRUE)
sd_fu <- sd(clean_data$FU_time, na.rm = TRUE)
```

```{r echo=FALSE}
# Uncomment this section to examine whether cohorts vary on key variables

# # MMSE
# mmse_model <- aov(mmse ~ cohort, data = clean_data)
# summary(mmse_model)
#
# # Age baseline
# age_model <- aov(age_baseline ~ cohort, data = clean_data)
# summary(age_model)
#
# # Sex
# chisq.test(clean_data$sex, clean_data$cohort)
#
# # Education
# edu_model <- aov(edu ~ cohort, data = clean_data)
# summary(edu_model)
#
# # FU time
# fu_time_model <- aov(FU_time ~ cohort, data = clean_data)
# summary(fu_time_model)
#
# # FU observations
# fu_obs_model <- aov(FU_obs ~ cohort, data = clean_data)
# summary(fu_obs_model)
```

We included `r participants` participants: `r round(female_percentage, 2)`% female with a mean baseline age of `r round(mean_age, 2)` years (SD = `r round(sd_age, 2)`) and a mean follow-up time of `r round(mean_fu, 2)` years (SD = `r round(sd_fu, 2)`). The descriptive statistics are summarized by cohort in Table 3. At least two cohorts varied significantly on baseline age, sex, education, MMSE, follow-up time, and follow-up observations. 

```{r echo=FALSE}
# Create a new variable to represent average number of follow-ups
clean_data <- clean_data %>%
  mutate(FU_obs = rowMeans(select(., em_obs, hc_obs, fc_obs), na.rm = TRUE))

clean_data <- clean_data %>%
  mutate(Study = recode(cohort,
    "Betula" = "Umea Betula",
    "Cobra" = "Umea COBRA",
    "Oslo" = "Oslo LCBC",
    "WAHA" = "Barcelona WAHA"
  ))

summary_stats <- clean_data %>%
  # Create a custom order for the table
  mutate(Study = factor(Study, levels = c("Barcelona WAHA", "Oslo LCBC", "Umea Betula", "Umea COBRA"))) %>%
  arrange(Study)

summary_stats <- summary_stats %>%
  dplyr::group_by(Study) %>%
  dplyr::summarise(
    `**Participants**` = n_distinct(id),
    `**Baseline age (range)**` = paste0(
      round(mean(age_baseline, na.rm = T), digits = 1),
      " (", paste0(round(range(age_baseline, na.rm = T)),
        collapse = "-"
      ),
      ")"
    ),
    `**Sex (% F / % M)**` = paste0(
      c(
        round(mean(sex == "female", na.rm = T) * 100),
        round(mean(sex == "male", na.rm = T) * 100)
      ),
      collapse = "/"
    ),
    `**Education (range)**` = paste0(
      round(mean(edu, na.rm = T), digits = 1),
      " (", paste0(round(range(edu, na.rm = T)),
        collapse = "-"
      ),
      ")"
    ),
    `**MMSE (range)**` = paste0(
      round(mean(mmse, na.rm = T), digits = 1),
      " (", paste0(round(range(mmse, na.rm = T)),
        collapse = "-"
      ),
      ")"
    ),
    `**Mean follow-up time (range)**` = paste0(
      round(mean(FU_time, na.rm = T), digits = 1),
      " (", paste0(round(range(FU_time, na.rm = T)),
        collapse = "-"
      ),
      ")"
    ),
    `**Mean follow-up observations (range)**` = paste0(
      round(mean(FU_obs, na.rm = TRUE), digits = 1),
      " (",
      "2",
      "-",
      round(max(FU_obs, na.rm = TRUE)),
      ")"
    )
  )

# Create a data frame for the total line
total_line <- tibble::tibble(
  `Study` = "Total",
  `**Participants**` = n_distinct(clean_data$id),
  `**Baseline age (range)**` = paste0(
    round(mean(clean_data$age_baseline, na.rm = T), digits = 1),
    " (", paste0(round(range(clean_data$age_baseline, na.rm = T)),
      collapse = "-"
    ),
    ")"
  ),
  `**Sex (% F / % M)**` = paste0(
    c(
      round(mean(clean_data$sex == "female", na.rm = T) * 100),
      round(mean(clean_data$sex == "male", na.rm = T) * 100)
    ),
    collapse = "/"
  ),
  `**Education (range)**` = paste0(
    round(mean(clean_data$edu, na.rm = T), digits = 1),
    " (", paste0(round(range(clean_data$edu, na.rm = T)),
      collapse = "-"
    ),
    ")"
  ),
  `**MMSE (range)**` = paste0(
    round(mean(clean_data$mmse, na.rm = T), digits = 1),
    " (", paste0(round(range(clean_data$mmse, na.rm = T)),
      collapse = "-"
    ),
    ")"
  ),
  `**Mean follow-up time (range)**` = paste0(
    round(mean(clean_data$FU_time, na.rm = T), digits = 1),
    " (", paste0(round(range(clean_data$FU_time, na.rm = T)),
      collapse = "-"
    ),
    ")"
  ),
  `**Mean follow-up observations (range)**` = paste0(
    round(mean(clean_data$FU_obs, na.rm = TRUE), digits = 1),
    " (",
    "2",
    "-",
    round(max(clean_data$FU_obs, na.rm = TRUE)),
    ")"
  )
)

# Bind the total line data frame to the summary statistics table
summary_stats <- bind_rows(summary_stats, total_line)
```

__Table 3__: _Demographic characteristics_
```{r tab3, echo=FALSE}
knitr::kable(summary_stats, align = "c")
```

<small>_Note._ Abbreviations: MMSE Mini Mental State Examination, F female, M male, CI Confidence Interval</small>

We found that as age increased both memory and adjusted hippocampus volume decreased, though we also observed inter-individual differences in these relationships (Supplementary Figure 1). Both the LME model and GAMM fit to this data showed evidence of significant relationships between memory/adjusted hippocampus volume and age; however, there was evidence of a non-linear relationship in both models (Supplementary Table 1). We include further details in the supplementary materials.

__**Episodic Memory and Hippocampus Change:**__ 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Helper function to format p-values
format_p_value <- function(p) {
  if (p < 0.001) {
    return("<0.001") # Written like this because of HTML formatting
  } else {
    return(round(p, 3))
  }
}

# Helper function to perform regression and extract required statistics
get_regression_stats <- function(formula, data, vars_priority) {
  model <- lm(formula, data = data)
  summary_model <- summary(model)
  coefficients <- summary_model$coefficients

  # Initialize the variable index
  variable_index <- NULL

  # Check for each variable in priority order and break on the first match
  for (var in vars_priority) {
    if (var %in% rownames(coefficients)) {
      variable_index <- which(rownames(coefficients) == var)
      break
    }
  }

  # Check if a relevant variable was found
  if (is.null(variable_index)) {
    stop("None of the priority variables are in the model.")
  }

  coef <- coefficients[variable_index, 1]
  ci <- confint(model)[variable_index, ]
  p_value <- coefficients[variable_index, 4]
  t_stat <- coefficients[variable_index, 3]
  adj_r2 <- summary_model$adj.r.squared
  r2 <- summary_model$r.squared

  return(list(coef = coef, ci = ci, p_value = p_value, t_stat = t_stat, adj_r2 = adj_r2, r2 = r2))
}

# Set a list of priority variables
hc_vars <- c(
  "scale(hc_slopes)",
  "hc_slopes",
  "scale(adj_HIP_total_baseline)",
  "scale(adj_HIP_total_followup)"
)

edu_var <- c(
  "scale(edu)"
)

# Running the regression model
emhp_merged <- get_regression_stats(scale(memory_slopes) ~ scale(hc_slopes) + study + scale(age_baseline) + scale(mem_time) + sex + scale(edu), data = clean_data, hc_vars)
emhp_merged_unscaled <- get_regression_stats(memory_slopes ~ hc_slopes + study + age_baseline + mem_time + sex + edu, data = clean_data, hc_vars)

# Cross-sectional analysis at baseline
model_baseline <- get_regression_stats(scale(memory_baseline) ~ scale(adj_HIP_total_baseline) + study + scale(age_baseline) + scale(edu) + sex, data = clean_data, hc_vars)

# Cross-sectional analysis at follow-up
model_followup <- get_regression_stats(scale(memory_followup) ~ scale(adj_HIP_total_followup) + study + scale(age_followup) + scale(edu), data = clean_data, hc_vars)

# Education and memory
edu_baseline <- get_regression_stats(scale(memory_baseline) ~ study + scale(age_baseline) + scale(edu) + sex, data = clean_data, edu_var)
edu_followup <- get_regression_stats(scale(memory_followup) ~ study + scale(age_followup) + scale(edu), data = clean_data, edu_var)
edu_long <- get_regression_stats(scale(memory_slopes) ~ study + scale(age_baseline) + scale(mem_time) + sex + scale(edu), data = clean_data, edu_var)
```

Hippocampus annual change and memory annual change were significantly correlated (β = `r round(emhp_merged$coef, 3)`, _p_ = `r format_p_value(emhp_merged$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "emhp_merged", "adjusted_p_value"])`, r² = `r round(emhp_merged$r2, 3)`) (Figure 2). Cross-sectionally, hippocampal volume and memory were not significantly correlated at baseline (_p_ = `r format_p_value(model_baseline$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "model_baseline", "adjusted_p_value"])`) but were significantly correlated at follow-up (β = `r round(model_followup$coef, 3)`, _p_ = `r format_p_value(model_followup$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "model_followup", "adjusted_p_value"])`, r² = `r round(model_followup$r2, 3)`). Education was only a significant predictor of memory cross-sectionally at baseline (β = `r round(edu_baseline$coef, 3)`, _p_ = `r format_p_value(edu_baseline$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_baseline", "adjusted_p_value"])`, r² = `r round(edu_baseline$r2, 3)`) and at follow-up (β = `r round(edu_followup$coef, 3)`, _p_ = `r format(format_p_value(edu_followup$p_value), scientific = FALSE)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_followup", "adjusted_p_value"])`, r² = `r round(edu_followup$r2, 3)`) but not longitudinally (_p_ = `r format(format_p_value(edu_long$p_value), scientific = FALSE)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_long", "adjusted_p_value"])`). These results are summarized in Table 4. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot without colors by cohort
mem_hc_plot <- ggplot(clean_data, aes(y = res_mem_slopes, x = res_hc_slopes)) +
  geom_point() +
  geom_smooth(
    method = "lm",
    color = "#86B6F6",
    fill = "#B4D4FF"
  ) +
  labs(
    y = "Memory Annual Change",
    x = "Hippocampal Annual Change"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.background = element_rect(fill = "white", color = "white"), # Set background to white
    plot.margin = margin(t = 0.35, r = 0.35, b = 0.35, l = 0.35, unit = "cm"),
    axis.text.x = element_text(margin = margin(t = -5)),
    plot.title = element_text(hjust = 0.5)
  )
```

__Figure 2:__ _Memory Annual Change vs Hippocampal Annual Change_
```{r fig1, echo=FALSE, dpi = 400, fig.width=6, fig.height=4, message=FALSE, warning=FALSE}
mem_hc_plot
```

The association between memory annual change and hippocampus annual change, controlled for MRI scanner/cohort, with a linear model with a coefficient of `r format(round(emhp_merged_unscaled$coef, 4), scientific = FALSE)` (scaled: `r round(emhp_merged$coef, 4)`) and confidence intervals. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create the table
table_data_emhc <- data.frame(
  " " = c(
    "", "**Memory Annual Change**",
    "", "", "**Memory Annual Change**"
  ),
  "**Hippocampus Annual Change**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(emhp_merged$coef, 3), " (", round(emhp_merged$ci[1], 3), " to ", round(emhp_merged$ci[2], 3), ")"),
    "**Education Longitudinal**", "_Coefficient (95% CI)_",
    paste0(round(edu_long$coef, 3), " (", round(edu_long$ci[1], 3), " to ", round(edu_long$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "emhp_merged", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "edu_long", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(emhp_merged$adj_r2, 3),
    "", "_Adjusted R²_", round(edu_long$adj_r2, 3)
  ),
  "**Hippocampus Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(model_baseline$coef, 3), " (", round(model_baseline$ci[1], 3), " to ", round(model_baseline$ci[2], 3), ")"),
    "**Education Baseline**", "_Coefficient (95% CI)_",
    paste0(round(edu_baseline$coef, 3), " (", round(edu_baseline$ci[1], 3), " to ", round(edu_baseline$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "model_baseline", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "edu_baseline", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(model_baseline$adj_r2, 3),
    "", "_Adjusted R²_", round(edu_baseline$adj_r2, 3)
  ),
  "**Hippocampus Follow-up**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(model_followup$coef, 3), " (", round(model_followup$ci[1], 3), " to ", round(model_followup$ci[2], 3), ")"),
    "**Education Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(edu_followup$coef, 3), " (", round(edu_followup$ci[1], 3), " to ", round(edu_followup$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "model_followup", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "edu_followup", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(model_followup$adj_r2, 3),
    "", "_Adjusted R²_", round(edu_followup$adj_r2, 3)
  )
)
```

__Table 4__: _Linear Regressions with Hippocampal Annual Change, Education, and Memory Annual Change_
```{r tab4, echo=FALSE} 
# Format and print the table
kable(table_data_emhc, format = "markdown", col.names = c(" ", "**Hippocampus Annual Change**", " ", " ", "**Hippocampus Baseline**", " ", " ", "**Hippocampus Follow-up**", " ", " "))
```
Note. Models are adjusted for age, sex, education, follow-up time, and cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: CI Confidence Interval

```{r echo=FALSE}
# Calculate mean and SD for CR and BM variables
mean_cr <- mean(clean_data$reverse_cr, na.rm = TRUE)
sd_cr <- sd(clean_data$reverse_cr, na.rm = TRUE)
mean_bm <- mean(clean_data$reverse_bm, na.rm = TRUE)
sd_bm <- sd(clean_data$reverse_bm, na.rm = TRUE)

# Calculate the combined CR and BM value
clean_data <- clean_data %>%
  mutate(
    total_cr_or_bm = ifelse(reverse_bm > reverse_cr, reverse_bm, -reverse_cr),
    closer_to = ifelse(reverse_bm > reverse_cr, "BM", "CR"),
    distance_from_midline = ifelse(closer_to == "BM", (reverse_bm / reverse_cr), (-(reverse_cr / reverse_bm)))
  )

mean_distance_midline <- mean(clean_data$distance_from_midline, na.rm = TRUE)
sd_distance_midline <- sd(clean_data$distance_from_midline, na.rm = TRUE)

# Classify participants into CR and BM pathways based on some criteria
# (Replace with actual classification logic if different)
cr_pathway <- clean_data %>%
  filter(closer_to == "CR") %>%
  nrow()
bm_pathway <- clean_data %>%
  filter(closer_to == "BM") %>%
  nrow()
total_participants <- nrow(clean_data)

# Run analyses for validating the CR variable as a measure of CR
cr_var <- c(
  "scale(reverse_cr)"
)

# Running the regression model
cr_validation <- get_regression_stats(scale(memory_slopes) ~ scale(hc_slopes) + scale(reverse_cr) + study + scale(age_baseline) + scale(mem_time) + sex + scale(edu), data = clean_data, cr_var)

```

__**CR and BM Variables and Index:**__

We established CR (mean = `r round(mean_cr, 3)`, SD = `r round(sd_cr, 3)`) and BM (mean = `r round(mean_bm, 3)`, SD = `r round(sd_bm, 3)`) variables (Figure 3). In a model with hippocampal annual change included, we found that CR remained a  significant predictor of memory annual change (β = `r round(cr_validation$coef, 3)`, _p_ = `r format_p_value(cr_validation$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_validation", "adjusted_p_value"])`, r² = `r round(emhp_merged$r2, 3)`).

```{r echo=FALSE, warning=FALSE}
custom_colors_bm <- c("#E5E483", "#8ac926", "#4CCD99", "#5BBCFF", "#024CAA", "#00224D")
custom_colors_cr <- c("#f5d7b0", "#f5be27", "#fa8d00", "#bc507f", "#7E1891", "#1c0159")

cr_plot <- ggplot(clean_data, aes(y = res_mem_slopes_scaled, x = res_hc_slopes_scaled, color = (reverse_cr))) +
  geom_point(size = 2) +
  scale_color_gradientn(
    colors = custom_colors_cr,
    limits = c(0.55, 2.59),
    values = c(0, 1 / 3, 2 / 3, 1) # Values mapped to scale so both figures have the same scale
  ) +
  geom_abline(slope = -(abs(max(clean_data$res_mem_slopes_scaled))) / (abs(min(clean_data$res_hc_slopes_scaled))), intercept = 0, color = "grey", size = 1) +
  geom_abline(slope = 0, intercept = 0, color = "grey", size = 1) +
  geom_abline(slope = 10000, intercept = 0, color = "grey", size = 1) +
  labs(
    y = "Memory Annual Change",
    x = "Hippocampus Annual Change",
    color = "Cognitive Reserve"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  xlim(-1, 1) +
  ylim(-1, 1)

bm_plot <- ggplot(clean_data, aes(y = res_mem_slopes_scaled, x = res_hc_slopes_scaled, color = reverse_bm)) +
  geom_point(size = 2) +
  scale_color_gradientn(
    colors = custom_colors_bm,
    limits = c(0.5, 2.6),
    values = c(0, 1 / 3, 2 / 3, 1) # Values mapped to scale so both figures have the same scale
  ) +
  geom_abline(
    slope = (abs(max(clean_data$res_mem_slopes_scaled)) / abs(min(clean_data$res_hc_slopes_scaled))),
    intercept = 0,
    color = "grey",
    size = 1
  ) +
  geom_abline(slope = 0, intercept = 0, color = "grey", size = 1) +
  geom_abline(slope = 10000, intercept = 0, color = "grey", size = 1) +
  labs(
    y = "Memory Annual Change",
    x = "Hippocampus Annual Change",
    color = "Brain Maintenance"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  xlim(-1, 1) +
  ylim(-1, 1)

# Combine the plots into one figure
combined_crbm <- bm_plot + cr_plot + plot_layout(ncol = 2)
```

__Figure 3:__ _Brain Maintenance and Cognitive Reserve Plots_

```{r fig2, echo=FALSE, dpi = 400, fig.width=7, fig.height=4, message=FALSE, warning=FALSE}
combined_crbm
```

<small>(A) The association between hippocampus annual change and memory annual change, adjusted for cohort/MRI scanner, depicted with a diagonal line representing the BM line. The color code represents the level of BM for a given participant. (B) The same association between hippocampus annual change and memory annual change, adjusted for cohort/MRI scanner, this time depicted with a diagonal line representing the CR line. The color code represents the level of CR for a given participant.</small>

We then created the CR and BM index with participants classified as on the CR pathway (N = `r cr_pathway`/`r total_participants`) or the BM pathway (N = `r bm_pathway`/`r total_participants`) (Figure 4). The mean absolute index value was `r round(mean_distance_midline, 3)` (SD = `r round(sd_distance_midline, 3)`).

```{r echo=FALSE, warning=FALSE}
# Create breaks for the graph looks nice
breaks_values <- c(seq(-5, 5, by = 1), 0)
breaks_values <- sort(unique(breaks_values))

# index_plot <- ggplot(clean_data, aes(x = index, y = total_cr_or_bm, fill = closer_to)) +
#   geom_point(aes(color = closer_to), shape = 21, size = 3, position = position_jitter(width = 0.01)) +
#   scale_fill_manual(values = c("BM" = "#AD88C6", "CR" = "#A1D6B2")) +
#   scale_color_manual(values = c("BM" = "#000000", "CR" = "#000000")) +
#   # facet_zoom(xlim = c(-0.5, 0.5)) + # Add this for the plot to zoom in in a 2nd variation
#   scale_x_continuous(
#     name = "Index Value",
#     breaks = breaks_values,
#     labels = scales::label_number(accuracy = .1),
#     expand = c(0, 0),
#     limits = c(-2.2, 2.2) # Set x-axis limits explicitly
#   ) +
#   labs(
#     title = "Cognitive Reserve / Brain Maintenance Index",
#     y = "Total CR or BM",
#     fill = "Pathway",
#     color = "Pathway"
#   ) +
#   theme_minimal() +
#   theme(
#     axis.text.y = element_text(),
#     axis.ticks.y = element_line(),
#     panel.grid.major.y = element_line(),
#     panel.grid.minor.y = element_blank(),
#     legend.position = "bottom",
#     plot.margin = margin(t = 0.25, r = 0.25, b = 0.25, l = 0.25, unit = "cm"),
#     axis.text.x = element_text(margin = margin(t = -5)),
#     plot.title = element_text(hjust = 0.5),
#   )

index_plot <- ggplot(clean_data, aes(x = total_cr_or_bm, y = 0, fill = closer_to)) +
  annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#d7eaf5") +
  annotate("rect", xmin = -Inf, xmax = 0, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#ffcfcf") +
  annotate("text", x = -1.5, y = 1.9, label = "CR Pathway", color = "black", size = 4, hjust = 0.5) +
  annotate("text", x = 1.5, y = 1.9, label = "BM Pathway", color = "black", size = 4, hjust = 0.5) +
  geom_violin(trim = FALSE, scale = "width", width = 4, position = position_identity()) +
  geom_jitter(aes(color = closer_to), shape = 21, size = 2.5, width = 0, height = 1) +
  scale_fill_manual(values = c("BM" = "#91c4e3", "CR" = "#FFAAAA")) +
  scale_color_manual(values = c("BM" = "#000000", "CR" = "#000000")) +
  scale_x_continuous(
    name = NULL,
    breaks = breaks_values,
    labels = abs(breaks_values),
    expand = c(0, 0),
    limits = c(-2.7, 2.7)
  ) +
  scale_y_continuous(
    limits = c(0, 2.1),
    expand = c(0, 0)
  ) +
  labs(
    fill = NULL,
    y = "Density",
    color = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white", color = "white"),
    plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"),
    plot.title = element_text(hjust = 0.5)
  )
```

__Figure 4:__ _Index of Brain Maintenance and Cognitive Reserve_

```{r fig3, echo=FALSE, dpi = 400, fig.width=8, fig.height=2, message=FALSE, warning=FALSE}
index_plot
```

An index depicting participants as either on the BM or CR pathway. The x-axis represents the index value – the ratio of CR to BM for participants with more CR and the ratio of BM to CR for participants with more BM. The y-axis represents the total CR for those on the CR path or the total BM for those of the BM path.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Prep the data for visualization

# Divide the data into those on the CR pathway or BM pathway 
cr_data <- clean_data %>% 
  filter(closer_to == "CR")

# Create groups of high and low
cr_data$reverse_cr_group <- cut(cr_data$reverse_cr,
  breaks = quantile(cr_data$reverse_cr, probs = seq(0, 1, (1 / 2)), na.rm = TRUE),
  include.lowest = TRUE,
  labels = c("Low CR", "High CR")
)

# And for BM
bm_data <- clean_data %>% 
  filter(closer_to == "BM")

bm_data$reverse_bm_group <- cut(bm_data$reverse_bm,
  breaks = quantile(bm_data$reverse_bm, probs = seq(0, 1, (1 / 2)), na.rm = TRUE),
  include.lowest = TRUE,
  labels = c("Low BM", "High BM")
)

# Convert df to long
long_data_vis_cr <- cr_data%>%
  pivot_longer(
    cols = matches("^memory\\.\\d$|^age\\.\\d$|^adj_HIP_total\\.\\d$"),
    names_to = c(".value", "timepoint"),
    names_pattern = "(.*)(\\d$)"
  )

long_data_vis_bm <- bm_data%>%
  pivot_longer(
    cols = matches("^memory\\.\\d$|^age\\.\\d$|^adj_HIP_total\\.\\d$"),
    names_to = c(".value", "timepoint"),
    names_pattern = "(.*)(\\d$)"
  )

long_data_vis_cr <- long_data_vis_cr %>%
  rename(
    age = "age.",
    adj_HIP_total = "adj_HIP_total.",
    memory = "memory."
  ) %>% 
  filter(!is.na(age) & !is.na(memory)) 

long_data_vis_bm <- long_data_vis_bm %>%
  rename(
    age = "age.",
    adj_HIP_total = "adj_HIP_total.",
    memory = "memory."
  ) %>% 
  filter(!is.na(age) & !is.na(memory)) 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Add a function for getting the stats from LME models
get_mixed_effects_stats <- function(formula, data, vars_priority) {
  # Fit the mixed-effects model
  model <- lmer(formula, data = data)
  summary_model <- summary(model)
  coefficients <- summary_model$coefficients
  anova_results <- anova(model) # Get ANOVA results for p-values

  # Initialize the variable index
  variable_index <- NULL

  # Check for each variable in priority order and break on the first match
  for (var in vars_priority) {
    if (var %in% rownames(coefficients)) {
      variable_index <- which(rownames(coefficients) == var)
      break
    }
  }

  # Check if a relevant variable was found
  if (is.null(variable_index)) {
    stop("None of the priority variables are in the model.")
  }

  coef <- coefficients[variable_index, 1]
  ci <- confint(model, parm = rownames(coefficients)[variable_index])

  # Extract p-value from ANOVA table
  p_value <- anova_results[which(rownames(anova_results) == rownames(coefficients)[variable_index]), "Pr(>F)"]
  r2 <- as.data.frame(VarCorr(model))$vcov[1]

  return(list(coef = coef, ci = ci, p_value = p_value, r2 = r2))
}

# Set a list of priority variables
vars <- c(
  "scale(age):scale(reverse_cr)",
  "scale(age):scale(reverse_bm)"
)

cr_age_lme <- get_mixed_effects_stats(
  scale(memory) ~ scale(age) * scale(reverse_cr) + study + scale(age_baseline) +
    scale(mem_time) + sex + scale(edu) + (1 | id),
  long_data_vis_cr,
  vars
)

bm_age_lme <- get_mixed_effects_stats(
  scale(memory) ~ scale(age) * scale(reverse_bm) + study + scale(age_baseline) +
    scale(mem_time) + sex + scale(edu) + (1 | id),
  long_data_vis_bm,
  vars
)
```

We then found a significant interaction between CR and age (β = `r round(cr_age_lme$coef, 3)`, _p_ = `r format_p_value(cr_age_lme$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_age_lme", "adjusted_p_value"])`, r² = `r round(cr_age_lme$r2, 3)`) and BM and age (β = `r round(bm_age_lme$coef, 3)`, _p_ = `r format_p_value(bm_age_lme$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_age_lme", "adjusted_p_value"])`, r² = `r round(bm_age_lme$r2, 3)`) on memory (Table 5). 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Create the table
table_cr_bm_age <- data.frame(
  " " = c(
    "", "**Memory**"
  ),
  "**Cognitive Reserve * Age**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(cr_age_lme$coef, 3), " (", round(cr_age_lme$ci[1], 3), " to ", round(cr_age_lme$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_age_lme", "adjusted_p_value"])
  ),
  " " = c(
    "_R²_", round(cr_age_lme$r2, 3)
  ),
  "**Brain Maintenance * Age**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(bm_age_lme$coef, 3), " (", round(bm_age_lme$ci[1], 3), " to ", round(bm_age_lme$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_age_lme", "adjusted_p_value"])
  ),
  " " = c(
    "_R²_", round(bm_age_lme$r2, 3)
  )
)
```

__Table 5__: _Linear Mixed Effects Models with Interactions between Age and CR / BM on Memory_
```{r tab5, echo=FALSE} 
# Format and print the table
kable(table_cr_bm_age, format = "markdown", col.names = c(" ", "**Brain Maintenance * Age**", " ", " ", "**Cognitive Reserve * Age**", " ", " "))
```
Note. Models are adjusted for age, sex, education, follow-up time, and cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: CI Confidence Interval

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Fit the mixed-effects model
model_res_lme <- lmer(memory ~ age * reverse_cr_group + (1 | id), data = long_data_vis_cr)

# Define a range of ages for prediction
age_seq <- seq(min(long_data_vis_cr$age), max(long_data_vis_cr$age), length.out = 100)

# Generate predictions for each CR group
# Divides the low and high CR groups into 100 age bands and,
# using the LME model from above, predicts their memory for each of the age bands
predictions <- do.call(rbind, lapply(unique(long_data_vis_cr$reverse_cr_group), function(res) {
  new_data <- data.frame(
    age = age_seq,
    reverse_cr_group = res,
    id = 1
  )
  pred <- predict(model_res_lme, newdata = new_data, se.fit = TRUE, re.form = NA)
  new_data$predicted <- pred$fit
  new_data$se <- pred$se.fit
  # Calculate 95% confidence intervals
  new_data$lower_bound <- new_data$predicted - 1.96 * new_data$se
  new_data$upper_bound <- new_data$predicted + 1.96 * new_data$se
  new_data
}))

# Define colors for each CR group
palette <- c("High CR" = "#7E1891", "Low CR" = "#f5be27")

# Create the plot
cr_age_plot <- ggplot() +
  geom_line(data = long_data_vis_cr, aes(x = age, y = memory, group = id), color = "lightgray", alpha = 0.5) + # Plot individual trajectories
  geom_point(data = long_data_vis_cr, aes(x = age, y = memory), color = "gray", size = 1) + # Add scatter points
  geom_ribbon(data = predictions, aes(x = age, ymin = lower_bound, ymax = upper_bound, fill = as.factor(reverse_cr_group)), alpha = 0.3) + # Add CIs
  geom_line(data = predictions, aes(x = age, y = predicted, color = as.factor(reverse_cr_group)), size = 1.2) + # Plot predicted lines for each group
  scale_color_manual(values = palette) +
  scale_fill_manual(values = palette) + # Ensure fill colors match line colors
  labs(x = "Age", y = "Memory", color = "Group", fill = "Group", title = "CR Pathway") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Fit the mixed-effects model
model_res_lme <- lmer(memory ~ age * reverse_bm_group + (1 | id), data = long_data_vis_bm)

# Define a range of ages for prediction
age_seq <- seq(min(long_data_vis_bm$age), max(long_data_vis_bm$age), length.out = 100)

# Generate predictions for each BM group
predictions <- do.call(rbind, lapply(unique(long_data_vis_bm$reverse_bm_group), function(res) {
  new_data <- data.frame(
    age = age_seq,
    reverse_bm_group = res,
    id = 1
  )
  pred <- predict(model_res_lme, newdata = new_data, se.fit = TRUE, re.form = NA)
  new_data$predicted <- pred$fit
  new_data$se <- pred$se.fit
  # Calculate 95% confidence intervals
  new_data$lower_bound <- new_data$predicted - 1.96 * new_data$se
  new_data$upper_bound <- new_data$predicted + 1.96 * new_data$se
  new_data
}))

# Define colors for each BM group
palette <- c("High BM" = "#1f77b4", "Low BM" = "#8ac926")

# Create the plot
bm_age_plot <- ggplot() +
  geom_line(data = long_data_vis_bm, aes(x = age, y = memory, group = id), color = "lightgray", alpha = 0.5) + # Plot individual trajectories
  geom_point(data = long_data_vis_bm, aes(x = age, y = memory), color = "gray", size = 1) + # Add scatter points
  geom_ribbon(data = predictions, aes(x = age, ymin = lower_bound, ymax = upper_bound, fill = as.factor(reverse_bm_group)), alpha = 0.3) + # Add CIs
  geom_line(data = predictions, aes(x = age, y = predicted, color = as.factor(reverse_bm_group)), size = 1.2) + # Plot predicted lines for each group
  scale_color_manual(values = palette) +
  scale_fill_manual(values = palette) + # Ensure fill colors match line colors
  labs(x = "Age", y = "Memory", color = "Group", fill = "Group", title = "BM Pathway") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Combine the plots into one figure
combined_plot <- bm_age_plot + cr_age_plot + plot_layout(ncol = 2)
```
__Figure 5:__ _The Relationship Between Memory and Age Depicted by CR or BM Group_

```{r fig4, echo=FALSE, dpi = 400, fig.width=6, fig.height=4, message=FALSE, warning=FALSE}
combined_plot
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
high_cr <- long_data_vis_cr %>%
  filter(reverse_cr_group == "High CR")

low_cr <- long_data_vis_cr %>%
  filter(reverse_cr_group == "Low CR")

high_bm <- long_data_vis_bm %>%
  filter(reverse_bm_group == "High BM")

low_bm <- long_data_vis_bm %>%
  filter(reverse_bm_group == "Low BM")

vars <- c("scale(age)")

vars_unscale <- c("age")

suppressMessages({
  high_cr_age <- get_mixed_effects_stats(scale(memory) ~ scale(age) + (1 | id), data = high_cr, vars)
  high_cr_age_unscaled <- get_mixed_effects_stats((memory) ~ age + (1 | id), data = high_cr, vars_unscale)

  low_cr_age <- get_mixed_effects_stats(scale(memory) ~ scale(age) + (1 | id), data = low_cr, vars)
  low_cr_age_unscaled <- get_mixed_effects_stats((memory) ~ (age) + (1 | id), data = low_cr, vars_unscale)

  high_bm_age <- get_mixed_effects_stats(scale(memory) ~ scale(age) + (1 | id), data = high_bm, vars)
  high_bm_age_unscaled <- get_mixed_effects_stats((memory) ~ age + (1 | id), data = high_bm, vars_unscale)

  low_bm_age <- get_mixed_effects_stats(scale(memory) ~ scale(age) + (1 | id), data = low_bm, vars)
  low_bm_age_unscaled <- get_mixed_effects_stats((memory) ~ (age) + (1 | id), data = low_bm, vars_unscale)
})
```
<small>(A) The association between memory and age depicted divided into two groups: high BM (blue) and low BM (green). There are linear mixed effects models for each group with the high BM group displaying a coefficient of `r round(high_bm_age_unscaled$coef, 3)` (scaled: `r round(high_bm_age$coef, 3)`) and the low BM group displaying a coefficient of `r round(low_bm_age_unscaled$coef, 3)` (scaled: `r round(low_bm_age$coef, 3)`) and confidence intervals. (B) The association between memory and age depicted divided into high CR (purple) and low CR (orange) with the high CR group displaying a coefficient of `r round(high_cr_age_unscaled$coef, 3)` (scaled: `r round(high_cr_age$coef, 3)`) and the low CR group displaying a coefficient of `r round(low_cr_age_unscaled$coef, 3)` (scaled: `r round(low_cr_age$coef, 3)`) and confidence intervals.</small>

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Education and CR / BM
vars <- c(
  "scale(reverse_cr)",
  "scale(reverse_bm)"
)

int_vars <- c(
  "scale(reverse_cr):scale(edu)",
  "scale(reverse_bm):scale(edu)"
)

edu_cr <- get_regression_stats(scale(edu) ~ scale(reverse_cr) + study, data = clean_data, vars)
edu_bm <- get_regression_stats(scale(edu) ~ scale(reverse_bm) + study, data = clean_data, vars)

edu_cr_int <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_cr) * scale(edu) + study, data = clean_data, int_vars)
edu_bm_int <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_bm) * scale(edu) + study, data = clean_data, int_vars)
```

__**BM / CR and Education:**__

There were no significant relationships between CR (_p_ = `r format_p_value(edu_cr$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_cr", "adjusted_p_value"])`) or BM (_p_ = `r format_p_value(edu_bm$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_bm", "adjusted_p_value"])`) and education. There was no significant interaction BM (_p_ = `r format_p_value(edu_bm_int$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_bm_int", "adjusted_p_value"])`) and education regressed on memory annual change. There was limited evidence of a significant interaction between CR and education regressed on memory annual change; however, this did not survive correction for multiple comparisons (β = `r round(edu_cr_int$coef, 3)`, _p_ = `r format_p_value(edu_cr_int$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "edu_cr_int", "adjusted_p_value"])`, r² = `r round(edu_cr_int$r2, 3)`). All education and CR/BM results are summarized in Table 6. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a BM / CR & education table
edu_table_data <- data.frame(
  " " = c(
    "", "**Education**",
    "", "",
    "**Memory Annual Change**"
  ),
  "**Brain Maintenance**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(edu_bm$coef, 3), " (", round(edu_bm$ci[1], 3), " to ", round(edu_bm$ci[2], 3), ")"),
    "**Brain Maintenance * Education**", "_Coefficient (95% CI)_",
    paste0(round(edu_bm_int$coef, 3), " (", round(edu_bm_int$ci[1], 3), " to ", round(edu_bm_int$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_",
    format_p_value(p_value_df[p_value_df$model_name == "edu_bm", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "edu_bm_int", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(edu_bm$adj_r2, 3),
    "", "_Adjusted R²_", round(edu_bm_int$adj_r2, 3)
  ),
  "**Cognitive Reserve**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(edu_cr$coef, 3), " (", round(edu_cr$ci[1], 3), " to ", round(edu_cr$ci[2], 3), ")"),
    "**Cognitive Reserve * Education**", "_Coefficient (95% CI)_",
    paste0(round(edu_cr_int$coef, 3), " (", round(edu_cr_int$ci[1], 3), " to ", round(edu_cr_int$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_",
    format_p_value(p_value_df[p_value_df$model_name == "edu_cr", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "edu_cr_int", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(edu_cr$adj_r2, 3),
    "", "_Adjusted R²_", round(edu_cr_int$adj_r2, 3)
  )
)
```

__Table 6__: _Linear Regressions with Cognitive Reserve / Brain Maintenance and Education_
```{r tab6, echo=FALSE} 
# Format and print the table
kable(edu_table_data, format = "markdown", col.names = c(" ", "**Cognitive Reserve**", " ", " ", "**Brain Maintenance**", " ", " "))
```

Note. Models are adjusted for cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: CI Confidence Interval

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Memory and function connectivity regs for BM

# Set the priority variables
priority_var_bm <- c(
  "scale(reverse_bm)"
)

priority_var_int_bm <- c(
  "scale(reverse_bm):scale(dmn_slopes)",
  "scale(reverse_bm):scale(ec_slopes)",
  "scale(reverse_bm):scale(sn_slopes)",
  "scale(reverse_bm):scale(dmn_baseline)",
  "scale(reverse_bm):scale(ec_baseline)",
  "scale(reverse_bm):scale(sn_baseline)",
  "scale(reverse_bm):scale(dmn_followup)",
  "scale(reverse_bm):scale(ec_followup)",
  "scale(reverse_bm):scale(sn_followup)"
)


bm_dmn_stats <- get_regression_stats(scale(dmn_slopes) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_dmn <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_bm) * scale(dmn_slopes) + study, data = clean_data, priority_var_int_bm)

bm_ecn_stats <- get_regression_stats(scale(ec_slopes) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_ec <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_bm) * scale(ec_slopes) + study, data = clean_data, priority_var_int_bm)

bm_sn_stats <- get_regression_stats(scale(sn_slopes) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_sn <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_bm) * scale(sn_slopes) + study, data = clean_data, priority_var_int_bm)

# Cross-sectional analyses
bm_dmn_stats_bl <- get_regression_stats(scale(dmn_baseline) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_dmn_stats_fu <- get_regression_stats(scale(dmn_followup) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_dmn_bl <- get_regression_stats(scale(memory_baseline) ~ scale(reverse_bm) * scale(dmn_baseline) + study, data = clean_data, priority_var_int_bm)
bm_dmn_fu <- get_regression_stats(scale(memory_followup) ~ scale(reverse_bm) * scale(dmn_baseline) + study, data = clean_data, priority_var_int_bm)

bm_ecn_stats_bl <- get_regression_stats(scale(ec_baseline) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_ecn_stats_fu <- get_regression_stats(scale(ec_followup) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_ec_bl <- get_regression_stats(scale(memory_baseline) ~ scale(reverse_bm) * scale(ec_baseline) + study, data = clean_data, priority_var_int_bm)
bm_ec_fu <- get_regression_stats(scale(memory_followup) ~ scale(reverse_bm) * scale(ec_baseline) + study, data = clean_data, priority_var_int_bm)

bm_sn_stats_bl <- get_regression_stats(scale(sn_baseline) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_sn_stats_fu <- get_regression_stats(scale(sn_followup) ~ scale(reverse_bm) + study, data = clean_data, priority_var_bm)
bm_sn_bl <- get_regression_stats(scale(memory_baseline) ~ scale(reverse_bm) * scale(sn_baseline) + study, data = clean_data, priority_var_int_bm)
bm_sn_fu <- get_regression_stats(scale(memory_followup) ~ scale(reverse_bm) * scale(sn_baseline) + study, data = clean_data, priority_var_int_bm)
```

__**BM / CR and Functional Connectivity:**__ 

There was no evidence of a significant association between BM and DMN annual change (_p_ = `r format_p_value(bm_dmn_stats$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_stats", "adjusted_p_value"])`) nor ECN annual change (_p_ = `r format_p_value(bm_ecn_stats$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_ecn_stats", "adjusted_p_value"])`). There was limited evidence of a significant association between BM and anterior SN annual change; however, this association did not survive correction for multiple comparisons (β = `r round(bm_sn_stats$coef, 3)`, _p_ = `r format_p_value(bm_sn_stats$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_sn_stats", "adjusted_p_value"])`, r² = `r round(bm_sn_stats$r2, 3)`). Additionally, there were no significant interactions between BM and both ECN annual change and anterior SN annual change regressed on memory annual change (ECN _p_ = `r format_p_value(bm_ec$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_ec", "adjusted_p_value"])`; anterior SN _p_ = `r format_p_value(bm_sn$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_sn", "adjusted_p_value"])`). Though we found some evidence of an interaction between BM and DMN annual change regressed on memory annual change, the interaction did not survive correction for multiple comparisons (BM/DMN β = `r round(bm_dmn$coef, 3)`, _p_ = `r format_p_value(bm_dmn$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_dmn", "adjusted_p_value"])`, r² = `r round(bm_dmn$r2, 3)`). Cross-sectionally at baseline and follow-up, there were no significant associations between BM and functional connectivity nor any significant interactions between BM and functional connectivity regressed on memory (Supplementary Table 2). 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Memory and function connectivity regs for CR

# Set the priority variables
priority_var_cr <- c(
  "scale(reverse_cr)"
)

priority_var_int_cr <- c(
  "scale(reverse_cr):scale(dmn_slopes)",
  "scale(reverse_cr):scale(ec_slopes)",
  "scale(reverse_cr):scale(sn_slopes)",
  "scale(reverse_cr):scale(dmn_baseline)",
  "scale(reverse_cr):scale(ec_baseline)",
  "scale(reverse_cr):scale(sn_baseline)",
  "scale(reverse_cr):scale(dmn_followup)",
  "scale(reverse_cr):scale(ec_followup)",
  "scale(reverse_cr):scale(sn_followup)"
)

cr_dmn_stats <- get_regression_stats(scale(dmn_slopes) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_dmn <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_cr) * scale(dmn_slopes) + study, data = clean_data, priority_var_int_cr)

cr_ecn_stats <- get_regression_stats(scale(ec_slopes) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_ec <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_cr) * scale(ec_slopes) + study, data = clean_data, priority_var_int_cr)

cr_sn_stats <- get_regression_stats(scale(sn_slopes) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_sn <- get_regression_stats(scale(memory_slopes) ~ scale(reverse_cr) * scale(sn_slopes) + study, data = clean_data, priority_var_int_cr)

# Cross-sectional analyses
cr_dmn_stats_bl <- get_regression_stats(scale(dmn_baseline) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_dmn_stats_fu <- get_regression_stats(scale(dmn_followup) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_dmn_bl <- get_regression_stats(scale(memory_baseline) ~ scale(reverse_cr) * scale(dmn_baseline) + study, data = clean_data, priority_var_int_cr)
cr_dmn_fu <- get_regression_stats(scale(memory_followup) ~ scale(reverse_cr) * scale(dmn_baseline) + study, data = clean_data, priority_var_int_cr)

cr_ecn_stats_bl <- get_regression_stats(scale(ec_baseline) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_ecn_stats_fu <- get_regression_stats(scale(ec_followup) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_ec_bl <- get_regression_stats(scale(memory_baseline) ~ scale(reverse_cr) * scale(ec_baseline) + study, data = clean_data, priority_var_int_cr)
cr_ec_fu <- get_regression_stats(scale(memory_followup) ~ scale(reverse_cr) * scale(ec_baseline) + study, data = clean_data, priority_var_int_cr)

cr_sn_stats_bl <- get_regression_stats(scale(sn_baseline) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_sn_stats_fu <- get_regression_stats(scale(sn_followup) ~ scale(reverse_cr) + study, data = clean_data, priority_var_cr)
cr_sn_bl <- get_regression_stats(scale(memory_baseline) ~ scale(reverse_cr) * scale(sn_baseline) + study, data = clean_data, priority_var_int_cr)
cr_sn_fu <- get_regression_stats(scale(memory_followup) ~ scale(reverse_cr) * scale(sn_baseline) + study, data = clean_data, priority_var_int_cr)
```

There was no evidence of a significant association between CR and DMN annual change (_p_ = `r format_p_value(cr_dmn_stats$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_stats", "adjusted_p_value"])`), ECN annual change (_p_ = `r format_p_value(cr_ecn_stats$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_ecn_stats", "adjusted_p_value"])`), nor anterior SN annual change (_p_ = `r format_p_value(cr_sn_stats$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_sn_stats", "adjusted_p_value"])`). Nor was there a significant interaction between CR and DMN annual change regressed on memory annual change (_p_ = `r format_p_value(cr_dmn$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_dmn", "adjusted_p_value"])`). However, there was a significant interaction between CR and both ECN annual change (β = `r round(cr_ec$coef, 3)`, _p_ = `r format(format_p_value(cr_ec$p_value), scientific = FALSE)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_ec", "adjusted_p_value"])`, r² = `r round(cr_ec$r2, 3)`) and anterior SN annual change (β = `r round(cr_sn$coef, 3)`, _p_ = `r format_p_value(cr_sn$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_sn", "adjusted_p_value"])`, r² = `r round(cr_sn$r2, 3)`) regressed on memory annual change (Figure 6). All functional connectivity results are summarized in Table 7. Cross-sectionally at baseline and follow-up, there were no significant associations between CR and functional connectivity nor any significant interactions between CR and functional connectivity regressed on memory (Supplementary Table 3). 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Divide the continuous variable into groups to be able to better visualize the data
clean_data$reverse_cr_group <- cut(clean_data$reverse_cr,
  breaks = quantile(clean_data$reverse_cr, probs = seq(0, 1, (1 / 2)), na.rm = TRUE),
  include.lowest = TRUE,
  labels = c("Low CR", "High CR")
)

ec_int <- ggplot(clean_data, aes(x = res_ec_slopes, y = res_mem_slopes, color = reverse_cr_group)) +
  geom_point() +
  geom_smooth(method = "lm", aes(fill = reverse_cr_group), alpha = 0.3) +
  labs(
    y = "Memory Annual Change",
    x = "ECN Annual Change",
    title = "Executive Control Network",
    color = "Group",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom", # Move legend to the bottom
    plot.title = element_text(hjust = 0.5)
  )

sn_int <- ggplot(clean_data, aes(x = res_sn_slopes, y = res_mem_slopes, color = reverse_cr_group)) +
  geom_point() +
  geom_smooth(method = "lm", aes(fill = reverse_cr_group), alpha = 0.3) +
  labs(
    y = "Memory Annual Change",
    x = "Anterior SN Annual Change",
    title = "Anterior Salience Network",
    color = "Group",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5)
  )

# Combine the plots into one figure
combined_plot <- sn_int + ec_int + plot_layout(ncol = 2)
```

__Figure 6:__ _The Interaction Between Cognitive Reserve, Functional Connectivity Change, and Memory Change_

```{r fig5, echo=FALSE, dpi = 400, fig.width=7, fig.height=4, message=FALSE, warning=FALSE}
combined_plot
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
high_cr <- clean_data %>%
  filter(reverse_cr_group == "High CR")

low_cr <- clean_data %>%
  filter(reverse_cr_group == "Low CR")

vars <- c(
  "scale(dmn_slopes)",
  "scale(sn_slopes)",
  "scale(ec_slopes)"
)

vars_unscale <- c(
  "dmn_slopes",
  "sn_slopes",
  "ec_slopes"
)

high_cr_sn <- get_regression_stats(scale(memory_slopes) ~ scale(sn_slopes) + study, data = high_cr, vars)
high_cr_sn_unscaled <- get_regression_stats((memory_slopes) ~ sn_slopes + study, data = high_cr, vars_unscale)

low_cr_sn <- get_regression_stats(scale(memory_slopes) ~ scale(sn_slopes) + study, data = low_cr, vars)
low_cr_sn_unscaled <- get_regression_stats((memory_slopes) ~ (sn_slopes) + study, data = low_cr, vars_unscale)

high_cr_ec <- get_regression_stats(scale(memory_slopes) ~ scale(ec_slopes) + study, data = high_cr, vars)
high_cr_ec_unscaled <- get_regression_stats((memory_slopes) ~ (ec_slopes) + study, data = high_cr, vars_unscale)

low_cr_ec <- get_regression_stats(scale(memory_slopes) ~ scale(ec_slopes) + study, data = low_cr, vars)
low_cr_ec_unscaled <- get_regression_stats((memory_slopes) ~ (ec_slopes) + study, data = low_cr, vars_unscale)
```

<small>(A) The association between anterior SN annual change and memory annual change, adjusted for cohort/MRI scanner, depicted divided into two groups: high CR (blue) and low CR (orange). There are linear models for each group with the high CR group displaying a coefficient of `r round(high_cr_sn_unscaled$coef, 3)` (scaled: `r round(high_cr_sn$coef, 3)`) and the low CR group displaying a coefficient of `r round(low_cr_sn_unscaled$coef, 3)` (scaled: `r round(low_cr_sn$coef, 3)`) and confidence intervals. (B) The association between ECN annual change and memory annual change, adjusted for cohort/MRI scanner, depicted divided into the two groups with the high CR group displaying a coefficient of `r round(high_cr_ec_unscaled$coef, 3)` (scaled: `r round(high_cr_ec$coef, 3)`) and the low CR group displaying a coefficient of `r round(low_cr_ec_unscaled$coef, 3)` (scaled: `r round(low_cr_ec$coef, 3)`). Abbreviations: ECN Executive Control Network, SN Salience Network, CR Cognitive Reserve, BM Brain Maintenance</small>

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create the table
table_data <- data.frame(
  " " = c(
    "", "**DMN Annual Change**",
    "**ECN Annual Change**", "**SN Annual Change**", "", "",
    "**Memory Annual Change**", "",
    "", "**Memory Annual Change**",
    "", "", "**Memory Annual Change**"
  ),
  "**Cognitive Reserve**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(cr_dmn_stats$coef, 3), " (", round(cr_dmn_stats$ci[1], 3), " to ", round(cr_dmn_stats$ci[2], 3), ")"),
    paste0(round(cr_ecn_stats$coef, 3), " (", round(cr_ecn_stats$ci[1], 3), " to ", round(cr_ecn_stats$ci[2], 3), ")"),
    paste0(round(cr_sn_stats$coef, 3), " (", round(cr_sn_stats$ci[1], 3), " to ", round(cr_sn_stats$ci[2], 3), ")"),
    "**Cognitive Reserve * DMN Annual Change**", "_Coefficient (95% CI)_",
    paste0(round(cr_dmn$coef, 3), " (", round(cr_dmn$ci[1], 3), " to ", round(cr_dmn$ci[2], 3), ")"),
    "**Cognitive Reserve * ECN Annual Change**", "_Coefficient (95% CI)_",
    paste0(round(cr_ec$coef, 3), " (", round(cr_ec$ci[1], 3), " to ", round(cr_ec$ci[2], 3), ")"),
    "**Cognitive Reserve * SN Annual Change**", "_Coefficient (95% CI)_",
    paste0(round(cr_sn$coef, 3), " (", round(cr_sn$ci[1], 3), " to ", round(cr_sn$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_",
    format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_stats", "adjusted_p_value"]),
    format_p_value(p_value_df[p_value_df$model_name == "cr_ecn_stats", "adjusted_p_value"]),
    format_p_value(p_value_df[p_value_df$model_name == "cr_sn_stats", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_dmn", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_ec", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_sn", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(cr_dmn_stats$adj_r2, 3), round(cr_ecn_stats$adj_r2, 3), round(cr_sn_stats$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_dmn$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_ec$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_sn$adj_r2, 3)
  ),
  "**Brain Maintenance**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(bm_dmn_stats$coef, 3), " (", round(bm_dmn_stats$ci[1], 3), " to ", round(bm_dmn_stats$ci[2], 3), ")"),
    paste0(round(bm_ecn_stats$coef, 3), " (", round(bm_ecn_stats$ci[1], 3), " to ", round(bm_ecn_stats$ci[2], 3), ")"),
    paste0(round(bm_sn_stats$coef, 3), " (", round(bm_sn_stats$ci[1], 3), " to ", round(bm_sn_stats$ci[2], 3), ")"),
    "**Brain Maintenance * DMN Annual Change**", "_Coefficient (95% CI)_",
    paste0(round(bm_dmn$coef, 3), " (", round(bm_dmn$ci[1], 3), " to ", round(bm_dmn$ci[2], 3), ")"),
    "**Brain Maintenance * ECN Annual Change**", "_Coefficient (95% CI)_",
    paste0(round(bm_ec$coef, 3), " (", round(bm_ec$ci[1], 3), " to ", round(bm_ec$ci[2], 3), ")"),
    "**Brain Maintenance * SN Annual Change**", "_Coefficient (95% CI)_",
    paste0(round(bm_sn$coef, 3), " (", round(bm_sn$ci[1], 3), " to ", round(bm_sn$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_",
    format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_stats", "adjusted_p_value"]),
    format_p_value(p_value_df[p_value_df$model_name == "bm_ecn_stats", "adjusted_p_value"]),
    format_p_value(p_value_df[p_value_df$model_name == "bm_sn_stats", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_dmn", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_ec", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_sn", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(bm_dmn_stats$adj_r2, 3), round(bm_ecn_stats$adj_r2, 3), round(bm_sn_stats$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_dmn$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_ec$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_sn$adj_r2, 3)
  )
)
```

__Table 7__: _Linear Regressions with Cognitive Reserve / Brain Maintenance and Functional Connectivity_
```{r tab7, echo=FALSE} 
# Format and print the table
kable(table_data, format = "markdown", col.names = c(" ", "**Cognitive Reserve**", " ", " ", "**Brain Maintenance**", " ", " "))
```

Note. Models are adjusted for cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: DMN Default Mode Network, ECN Executive Control Network, SN Salience Network, CI Confidence Interval

**Supplementary Materials**

**_LME and GAM Models for Memory / Adjusted Hippocampus Volume and Age_**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Convert df to long format
long_clean_data <- clean_data %>%
  pivot_longer(
    cols = matches("^memory\\.\\d$|^age\\.\\d$|^adj_HIP_total\\.\\d$"),
    names_to = c(".value", "timepoint"),
    names_pattern = "(.*)(\\d$)"
  )

long_clean_data <- long_clean_data %>%
  rename(
    age = "age.",
    adj_HIP_total = "adj_HIP_total.",
    memory = "memory."
  )

# Linear model for comparison with the gam model
mod_lme_mem <- lmer(memory ~ age + (1 | id) + study, data = long_clean_data)
aic_lm_mem <- AIC(mod_lme_mem)
bic_lm_mem <- BIC(mod_lme_mem)

mod_lme_hc <- lmer(adj_HIP_total ~ age + (1 | id) + study, data = long_clean_data)
aic_lm_hc <- AIC(mod_lme_hc)
bic_lm_hc <- BIC(mod_lme_hc)

# Get the regression stats
age <- c("scale(age)")

suppressMessages({
  mem_age_lme <- get_mixed_effects_stats(scale(memory) ~ scale(age) + (1 | id) + study, data = long_clean_data, age)
  hc_age_lme <- get_mixed_effects_stats(scale(adj_HIP_total) ~ scale(age) + (1 | id) + study, data = long_clean_data, age)
})
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Uncomment this code to evaluate and check GAMM assumptions

# # Loop through k values and print results
# for (k in 1:27) {
#   model_mem <- gamm4(memory ~ s(age, k = k+3) + study, data = long_clean_data, random = ~(1|id))
#
#   cat("Model with k =", k + 3, "\n")
#   cat("AIC:", AIC(model_mem$mer), "\n")
#   cat("BIC:", BIC(model_mem$mer), "\n")
#   cat("LogLik:", logLik(model_mem$mer), "\n")
#   cat("\n")
# }
#
# # Loop through k values and print results
# for (k in 1:27) {
#   model_hc <- gamm4(adj_HIP_total ~ s(age, k = k+3) + study, data = long_clean_data, random = ~(1|id))
#
#   cat("Model with k =", k + 3, "\n")
#   cat("AIC:", AIC(model_hc$mer), "\n")
#   cat("BIC:", BIC(model_hc$mer), "\n")
#   cat("LogLik:", logLik(model_hc$mer), "\n")
#   cat("\n")
# }
#
# gamm_memory <- gamm4(memory ~ s(age, k = 4) + study, data = long_clean_data, random = ~(1|id))
# gam.check(gamm_memory$gam)
#
# gamm_hc <- gamm4(adj_HIP_total ~ s(age, k = 5) + study, data = long_clean_data, random = ~(1|id))
# gam.check(gamm_hc$gam)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Function to calculate and store gamm model metrics
get_gamm_metrics <- function(k, df, response_var, lme_mod) {
  max_k <- min(k, length(unique(df$age)) - 1)

  if (max_k > 1) {
    # Dynamically use the response variable and correct random effect
    mod_em <- gamm4(as.formula(paste(response_var, "~ 1 + s(age, k =", max_k, ") +", make.names("study"))),
      data = df, random = ~ (1 | id)
    )

    # Calculate required metrics
    aic <- AIC(mod_em$mer)
    bic <- BIC(mod_em$mer)
    loglik <- logLik(mod_em$mer)

    # Extract EDF
    edf <- sum(summary(mod_em$gam)$edf)

    # Calculate p-value for the GAMM terms
    gamm_summary <- summary(mod_em$gam)
    p_value_gamm <- gamm_summary$s.pv

    # Likelihood ratio test
    lrt_stat <- -2 * (logLik(lme_mod) - loglik) # Comparing the lme to the gamm model
    lrt_df <- df.residual(lme_mod) - df.residual(mod_em$mer)
    p_value_lrt <- 1 - pchisq(lrt_stat, lrt_df)

    # Return a named list of metrics
    return(data.frame(
      k = max_k, AIC = aic, BIC = bic, LogLik = loglik,
      EDF = edf, LRT_df = lrt_df, "LRT test stat" = lrt_stat,
      "p_value_gamm" = as.numeric(p_value_gamm),
      "p_value_lrt" = as.numeric(p_value_lrt)
    ))
  } else {
    return(data.frame(
      k = max_k, AIC = NA, BIC = NA,
      LogLik = NA, EDF = NA, "LRT test stat" = NA,
      "p_value_gamm" = NA, "p_value_lrt" = NA
    ))
  }
}

# Function to fit and compare models over k range for different response variables
gamm_compare <- function(response_var, mod_lme) {
  compare_gamm <- do.call(rbind, lapply(k_range, function(k) {
    get_gamm_metrics(k, long_clean_data, response_var, mod_lme)
  }))
  return(compare_gamm)
}

# Run the comparison for memory and adjusted HIP
k_range <- 4 # Define k range for memory
gamm_compare_em_memory <- gamm_compare("memory", mod_lme_mem)

k_range <- 5 # Define k range for HC
gamm_compare_em_adj_hc <- gamm_compare("adj_HIP_total", mod_lme_hc)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
gamm_table <- data.frame(
  " " = c(
    "",
    "**GAMM**",
    "**LME Model**"
  ),
  "**Memory**" = c(
    "_AIC_",
    paste0(round(gamm_compare_em_memory$AIC, 3)),
    paste0(round(aic_lm_mem, 3))
  ),
  " " = c(
    "_BIC_",
    paste0(round(gamm_compare_em_memory$BIC, 3)),
    paste0(round(bic_lm_mem, 3))
  ),
  " " = c(
    "_Adjusted LRT p-value_",
    format_p_value(p_value_df[p_value_df$model_name == "gamm_memory_lft", "adjusted_p_value"]),
    " "
  ),
  "**Adjusted Hippocampus Volume**" = c(
    "_AIC_",
    paste0(round(gamm_compare_em_adj_hc$AIC, 3)),
    paste0(round(aic_lm_hc, 3))
  ),
  " " = c(
    "_BIC_",
    paste0(round(gamm_compare_em_adj_hc$BIC, 3)),
    paste0(round(bic_lm_hc, 3))
  ),
  " " = c(
    "_Adjusted LRT p-value_",
    format_p_value(p_value_df[p_value_df$model_name == "gamm_hc_lft", "adjusted_p_value"]),
    " "
  )
)
```

We compared a LME and GAMM fit to the memory/adjusted hippocampus volume and age data, while controlling for cohort/MRI scanner. For both models, we allowed a random intercept per participant. To assess model performance, we used the Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), and a likelihood ratio test (LRT) to compare the GAMM and LME models (Supplementary Table 1). The LRT evaluates whether the inclusion of non-linear terms enhances model fit, with a significant result indicating a preferable fit for the GAMM. The AIC and BIC also help identify the preferred model, with a difference of 10 or more favoring the model with lower values.

We initially fit the GAMM using a range of basis functions and compared their fit to the LME model. Despite variations in the GAMM fit, the conclusions regarding how the GAMM and the LME model compared did not change. In Supplementary Table 1, we fit the GAMM with four basis functions for the memory model and five basis functions for the adjusted hippocampus model as increasing the number beyond this point did not substantially change the AIC nor BIC values. The smooth age-term was significant in the memory GAMM (effective degrees of freedom (EDF) = `r round(gamm_compare_em_memory$EDF, 3)`, _p_-value = `r format_p_value(gamm_compare_em_memory$p_value_gamm)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "gamm_memory_gamm", "adjusted_p_value"])`) and the hippocampus GAMM (EDF = `r round(gamm_compare_em_adj_hc$EDF, 3)`, _p_-value = `r format_p_value(gamm_compare_em_adj_hc$p_value_gamm)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "gamm_hc_gamm", "adjusted_p_value"])`. For the LME models, both models had significant age coefficients (memory: β = `r round(mem_age_lme$coef, 3)`, _p_ = `r format(format_p_value(mem_age_lme$p_value), scientific = FALSE)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "mem_age_lme", "adjusted_p_value"])`; adjusted hippocampal volume β = `r round(hc_age_lme$coef, 3)`, _p_ = `r format(format_p_value(hc_age_lme$p_value), scientific = FALSE)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "hc_age_lme", "adjusted_p_value"])`). For memory, the LME and GAMM fits were similar; however, the GAMM fit was better according to AIC and LRT. For adjusted hippocampal volume, the GAMM fit was better according to AIC, BIC, and LRT.

__Supplementary Table 1__: _GAM vs LME Model_
```{r suptab1, echo=FALSE} 
# Format and print the table
kable(gamm_table, format = "markdown", col.names = c(" ", "**Memory**", " ", " ", "**Adjusted Hippocampus**", " ", " "))
```
Note. Models are adjusted for cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: GAMM Generalized Additive Mixed Model, LME Linear Mixed Effects, AIC Akaike Information Criterion, BIC Bayesian Information Criterion, LRT Likelihood Ratio Test

__Supplementary Figure 1__: _Memory and Hippocampus Volume Change with Age_
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Extract residuals
long_clean_data <- long_clean_data %>%
  filter(!is.na(age) & !is.na(memory) & !is.na(adj_HIP_total))

residual_model_age <- lm(age ~ study, data = long_clean_data)
residual_model_mem <- lm(memory ~ study, data = long_clean_data)
residual_model_hc <- lm(adj_HIP_total ~ study, data = long_clean_data)
long_clean_data$res_age <- residuals(residual_model_age)
long_clean_data$res_mem <- residuals(residual_model_mem)
long_clean_data$res_hc <- residuals(residual_model_hc)

# Define colors
color_palette <- c("indianred2", "steelblue2", "#86db49", "#F7CA18", "blue", "black")

# Create the plot age v memory
mem_age <- ggplot(long_clean_data, aes(x = age, y = memory, color = Study, group = id)) +
  geom_line(size = 0.4) +
  geom_point(size = 0.6) +
  scale_color_manual(values = color_palette) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 24, bs = "cs"), aes(group = 1), color = "grey", se = TRUE) +
  labs(x = "Age", y = "Episodic Memory", color = "Cohort") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Create the plot age v HC
hc_age <- ggplot(long_clean_data, aes(x = age, y = adj_HIP_total, color = Study, group = id)) +
  geom_line(size = 0.4) +
  geom_point(size = 0.6) +
  scale_color_manual(values = color_palette) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 24, bs = "cs"), aes(group = 1), color = "grey", se = TRUE) +
  labs(x = "Age", y = "Adjusted Hippocampal Volume", color = "Cohort") +
  theme_minimal() +
  theme(legend.position = "none") # Remove legend for the second plot

# Combine the plots and collect legend
hc_mem_age <- mem_age + hc_age + plot_layout(ncol = 2, guides = "collect") &
  theme(legend.position = "bottom")
```

```{r supfig1, echo=FALSE, dpi = 400, fig.width=7, fig.height=4, message=FALSE, warning=FALSE}
hc_mem_age
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
vars_unscale <- c(
  "age"
)

mem_age_unscaled <- get_mixed_effects_stats((memory) ~ age + (1 | id) + study, data = long_clean_data, vars_unscale)
```
<small>(A) The association between memory and age with individual trajectories and a generalized additive mixed model (EDF = `r round(gamm_compare_em_memory$EDF, 3)`) and confidence intervals. (B) The association between adjusted hippocampal volume and age with individual trajectories with a generalized additive mixed model (EDF = `r round(gamm_compare_em_adj_hc$EDF, 3)`) and confidence intervals.</small>

**_Cross-sectional BM / CR and Functional Connectivity_**

At baseline, there was no evidence of a significant association between BM and DMN connectivity (_p_ = `r format_p_value(bm_dmn_stats_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_stats_bl", "adjusted_p_value"])`), ECN connectivity (_p_ = `r format_p_value(bm_ecn_stats_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_ecn_stats_bl", "adjusted_p_value"])`), nor anterior SN connectivity (_p_ = `r format_p_value(bm_sn_stats_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_sn_stats_bl", "adjusted_p_value"])`). At follow-up, there was also no evidence of a significant association between BM and DMN connectivity (_p_ = `r format_p_value(bm_dmn_stats_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_stats_fu", "adjusted_p_value"])`), ECN connectivity (_p_ = `r format_p_value(bm_ecn_stats_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_ecn_stats_fu", "adjusted_p_value"])`), nor anterior SN connectivity (_p_ = `r format_p_value(bm_sn_stats_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_sn_stats_fu", "adjusted_p_value"])`). Regarding a cross-sectional interaction between functional connectivity and BM regressed on memory, at baseline, there was no evidence of a significant interaction between BM and DMN connectivity (_p_ = `r format_p_value(bm_dmn_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_bl", "adjusted_p_value"])`), ECN connectivity (_p_ = `r format_p_value(bm_ec_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_ec_bl", "adjusted_p_value"])`), nor anterior SN connectivity (_p_ = `r format_p_value(bm_sn_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_sn_bl", "adjusted_p_value"])`). Nor was there evidence of a significant interaction between BM and DMN connectivity (_p_ = `r format_p_value(bm_dmn_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_fu", "adjusted_p_value"])`), ECN connectivity (_p_ = `r format_p_value(bm_ec_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_ec_fu", "adjusted_p_value"])`), nor anterior SN connectivity (_p_ = `r format_p_value(bm_sn_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "bm_sn_fu", "adjusted_p_value"])`) regressed on memory at follow-up (Supplementary Table 2).

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create the table
cs_fc <- data.frame(
  " " = c(
    "", "**Brain Maintenance**",
    "", "", "**Brain Maintenance**",
    "", "", "**Memory Annual Change**",
    "", "", "**Memory Annual Change**"
  ),
  "**DMN Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(bm_dmn_stats_bl$coef, 3), " (", round(bm_dmn_stats_bl$ci[1], 3), " to ", round(bm_dmn_stats_bl$ci[2], 3), ")"),
    "**DMN Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(bm_dmn_stats_fu$coef, 3), " (", round(bm_dmn_stats_fu$ci[1], 3), " to ", round(bm_dmn_stats_fu$ci[2], 3), ")"),
    "**DMN Baseline * Brain Maintenance**", "_Coefficient (95% CI)_",
    paste0(round(bm_dmn_bl$coef, 3), " (", round(bm_dmn_bl$ci[1], 3), " to ", round(bm_dmn_bl$ci[2], 3), ")"),
    "**DMN Follow-up * Brain Maintenance**", "_Coefficient (95% CI)_",
    paste0(round(bm_dmn_fu$coef, 3), " (", round(bm_dmn_fu$ci[1], 3), " to ", round(bm_dmn_fu$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_stats_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_stats_fu", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_dmn_fu", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(bm_dmn_stats_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_dmn_stats_fu$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_dmn_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_dmn_fu$adj_r2, 3)
  ),
  "**ECN Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(bm_ecn_stats_bl$coef, 3), " (", round(bm_ecn_stats_bl$ci[1], 3), " to ", round(bm_ecn_stats_bl$ci[2], 3), ")"),
    "**ECN Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(bm_ecn_stats_fu$coef, 3), " (", round(bm_ecn_stats_fu$ci[1], 3), " to ", round(bm_ecn_stats_fu$ci[2], 3), ")"),
    "**ECN Baseline * Brain Maintenance **", "_Coefficient (95% CI)_",
    paste0(round(bm_ec_bl$coef, 3), " (", round(bm_ec_bl$ci[1], 3), " to ", round(bm_ec_bl$ci[2], 3), ")"),
    "**ECN Follow-up * Brain Maintenance**", "_Coefficient (95% CI)_",
    paste0(round(bm_ec_fu$coef, 3), " (", round(bm_ec_fu$ci[1], 3), " to ", round(bm_ec_fu$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_ecn_stats_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_ecn_stats_fu", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_ec_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_ec_fu", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(bm_ecn_stats_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_ecn_stats_fu$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_ec_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_ec_fu$adj_r2, 3)
  ),
  "**Anterior SN Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(bm_sn_stats_bl$coef, 3), " (", round(bm_sn_stats_bl$ci[1], 3), " to ", round(bm_sn_stats_bl$ci[2], 3), ")"),
    "**Anterior SN Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(bm_sn_stats_fu$coef, 3), " (", round(bm_sn_stats_fu$ci[1], 3), " to ", round(bm_sn_stats_fu$ci[2], 3), ")"),
    "**Anterior SN Baseline * Brain Maintenance**", "_Coefficient (95% CI)_",
    paste0(round(bm_sn_bl$coef, 3), " (", round(bm_sn_bl$ci[1], 3), " to ", round(bm_sn_bl$ci[2], 3), ")"),
    "**Anterior SN Follow-up * Brain Maintenance**", "_Coefficient (95% CI)_",
    paste0(round(bm_sn_fu$coef, 3), " (", round(bm_sn_fu$ci[1], 3), " to ", round(bm_sn_fu$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_sn_stats_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_sn_stats_fu", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_sn_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "bm_sn_fu", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(bm_sn_stats_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_sn_stats_fu$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_sn_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(bm_sn_fu$adj_r2, 3)
  )
)
```

__Supplementary Table 2__: _Cross-sectional Functional Connectivity, Brain Maintenance and Memory_
```{r suptab2, echo=FALSE} 
# Format and print the table
kable(cs_fc, format = "markdown", col.names = c(" ", "**DMN Baseline**", " ", " ", "**ECN Baseline**", " ", " ", "**Anterior SN Baseline**", " ", " "))
```
Note. Models are adjusted for cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: DMN Default Mode Network, ECN Executive Control Network, SN Salience Network, CI Confidence Interval

For CR, there was no evidence of a significant association between CR and DMN connectivity (_p_ = `r format_p_value(cr_dmn_stats_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_stats_bl", "adjusted_p_value"])`), ECN connectivity (_p_ = `r format_p_value(cr_ecn_stats_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_ecn_stats_bl", "adjusted_p_value"])`), nor anterior SN connectivity (_p_ = `r format_p_value(cr_sn_stats_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_sn_stats_bl", "adjusted_p_value"])`) at baseline. At follow-up, there was also no evidence of a significant association between CR and ECN connectivity (_p_ = `r format_p_value(cr_ecn_stats_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_ecn_stats_fu", "adjusted_p_value"])`) nor anterior SN connectivity (_p_ = `r format_p_value(cr_sn_stats_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_sn_stats_fu", "adjusted_p_value"])`). With the DMN, there was some evidence of a significant association between CR and DMN connectivity at follow-up (β = `r round(cr_dmn_stats_fu$coef, 3)`, _p_ = `r format_p_value(cr_dmn_stats_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_stats_fu", "adjusted_p_value"])`, r² = `r round(cr_dmn_stats_fu$r2, 3)`); however, this did not survive correction for multiple comparisons. Regarding a cross-sectional interaction between functional connectivity and CR, at baseline, there was no evidence of a significant interaction between CR and ECN connectivity (_p_ = `r format_p_value(cr_ec_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_ec_bl", "adjusted_p_value"])`) nor anterior SN connectivity (_p_ = `r format_p_value(cr_sn_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_sn_bl", "adjusted_p_value"])`) regressed on memory. Though there was some evidence of a significant interaction between CR and DMN regressed on memory at baseline (β = `r round(cr_dmn_bl$coef, 3)`, _p_ = `r format_p_value(cr_dmn_bl$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_bl", "adjusted_p_value"])`, r² = `r round(cr_dmn_bl$r2, 3)`) this did not survive correction for multiple comparisons. There was no evidence of a significant interaction between CR and DMN connectivity (_p_ = `r format_p_value(cr_dmn_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_fu", "adjusted_p_value"])`), ECN connectivity (_p_ = `r format_p_value(cr_ec_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_ec_fu", "adjusted_p_value"])`), nor anterior SN connectivity (_p_ = `r format_p_value(cr_sn_fu$p_value)`, _p_<sub>FDR</sub> = `r format_p_value(p_value_df[p_value_df$model_name == "cr_sn_fu", "adjusted_p_value"])`) regressed on memory at follow-up (Supplementary Table 3). 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create the table
cs_fc <- data.frame(
  " " = c(
    "", "**Cognitive Reserve**",
    "", "", "**Cognitive Reserve**",
    "", "", "**Memory Annual Change**",
    "", "", "**Memory Annual Change**"
  ),
  "**DMN Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(cr_dmn_stats_bl$coef, 3), " (", round(cr_dmn_stats_bl$ci[1], 3), " to ", round(cr_dmn_stats_bl$ci[2], 3), ")"),
    "**DMN Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(cr_dmn_stats_fu$coef, 3), " (", round(cr_dmn_stats_fu$ci[1], 3), " to ", round(cr_dmn_stats_fu$ci[2], 3), ")"),
    "**DMN Baseline * Cognitive Reserve**", "_Coefficient (95% CI)_",
    paste0(round(cr_dmn_bl$coef, 3), " (", round(cr_dmn_bl$ci[1], 3), " to ", round(cr_dmn_bl$ci[2], 3), ")"),
    "**DMN Follow-up * Cognitive Reserve**", "_Coefficient (95% CI)_",
    paste0(round(cr_dmn_fu$coef, 3), " (", round(cr_dmn_fu$ci[1], 3), " to ", round(cr_dmn_fu$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_stats_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_stats_fu", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_dmn_fu", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(cr_dmn_stats_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_dmn_stats_fu$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_dmn_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_dmn_fu$adj_r2, 3)
  ),
  "**ECN Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(cr_ecn_stats_bl$coef, 3), " (", round(cr_ecn_stats_bl$ci[1], 3), " to ", round(cr_ecn_stats_bl$ci[2], 3), ")"),
    "**ECN Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(cr_ecn_stats_fu$coef, 3), " (", round(cr_ecn_stats_fu$ci[1], 3), " to ", round(cr_ecn_stats_fu$ci[2], 3), ")"),
    "**ECN Baseline * Cognitive Reserve **", "_Coefficient (95% CI)_",
    paste0(round(cr_ec_bl$coef, 3), " (", round(cr_ec_bl$ci[1], 3), " to ", round(cr_ec_bl$ci[2], 3), ")"),
    "**ECN Follow-up * Cognitive Reserve**", "_Coefficient (95% CI)_",
    paste0(round(cr_ec_fu$coef, 3), " (", round(cr_ec_fu$ci[1], 3), " to ", round(cr_ec_fu$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_ecn_stats_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_ecn_stats_fu", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_ec_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_ec_fu", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(cr_ecn_stats_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_ecn_stats_fu$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_ec_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_ec_fu$adj_r2, 3)
  ),
  "**Anterior SN Baseline**" = c(
    "_Coefficient (95% CI)_",
    paste0(round(cr_sn_stats_bl$coef, 3), " (", round(cr_sn_stats_bl$ci[1], 3), " to ", round(cr_sn_stats_bl$ci[2], 3), ")"),
    "**Anterior SN Follow-up**", "_Coefficient (95% CI)_",
    paste0(round(cr_sn_stats_fu$coef, 3), " (", round(cr_sn_stats_fu$ci[1], 3), " to ", round(cr_sn_stats_fu$ci[2], 3), ")"),
    "**Anterior SN Baseline * Cognitive Reserve**", "_Coefficient (95% CI)_",
    paste0(round(cr_sn_bl$coef, 3), " (", round(cr_sn_bl$ci[1], 3), " to ", round(cr_sn_bl$ci[2], 3), ")"),
    "**Anterior SN Follow-up * Cognitive Reserve**", "_Coefficient (95% CI)_",
    paste0(round(cr_sn_fu$coef, 3), " (", round(cr_sn_fu$ci[1], 3), " to ", round(cr_sn_fu$ci[2], 3), ")")
  ),
  " " = c(
    "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_sn_stats_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_sn_stats_fu", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_sn_bl", "adjusted_p_value"]),
    "", "_Adjusted p-value_", format_p_value(p_value_df[p_value_df$model_name == "cr_sn_fu", "adjusted_p_value"])
  ),
  " " = c(
    "_Adjusted R²_", round(cr_sn_stats_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_sn_stats_fu$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_sn_bl$adj_r2, 3),
    "", "_Adjusted R²_", round(cr_sn_fu$adj_r2, 3)
  )
)
```

__Supplementary Table 3__: _Cross-sectional Functional Connectivity, Cognitive Reserve and Memory_
```{r suptab3, echo=FALSE} 
# Format and print the table
kable(cs_fc, format = "markdown", col.names = c(" ", "**DMN Baseline**", " ", " ", "**ECN Baseline**", " ", " ", "**Anterior SN Baseline**", " ", " "))
```
Note. Models are adjusted for cohort/MRI scanner. _p_-values are adjusted for multiple comparisons using the FDR. Abbreviations: DMN Default Mode Network, ECN Executive Control Network, SN Salience Network, CI Confidence Interval

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Correcting for p-values

# Memory, Hippocampus and Analyses
memory_hippocampus_edu_models <- list(
  emhp_merged, model_baseline, model_followup,
  edu_baseline, edu_followup, edu_long
)

# Education and CR/BM Analyses
education_models <- list(
  edu_cr, edu_bm, edu_cr_int, edu_bm_int
)

# CR and BM Sensitivity Analyses
cr_bm_models <- list(
  cr_validation, bm_age_lme, cr_age_lme
)

# Functional Connectivity and CR/BM Analyses Longitudinal
functional_connectivity_long <- list(
  bm_dmn_stats, bm_dmn, bm_ecn_stats, bm_ec, bm_sn_stats, bm_sn,
  cr_dmn_stats, cr_dmn, cr_ecn_stats, cr_ec, cr_sn_stats, cr_sn
)

# Functional Connectivity and CR/BM Analyses Cross-sectional
functional_connectivity_cs <- list(
  cr_dmn_bl, cr_dmn_fu, cr_dmn_stats_bl, cr_dmn_stats_fu,
  cr_ec_bl, cr_ec_fu, cr_ecn_stats_bl, cr_ecn_stats_fu,
  cr_sn_bl, cr_sn_fu, cr_sn_stats_bl, cr_sn_stats_fu,
  bm_dmn_bl, bm_dmn_fu, bm_dmn_stats_bl, bm_dmn_stats_fu,
  bm_ec_bl, bm_ec_fu, bm_ecn_stats_bl, bm_ecn_stats_fu,
  bm_sn_bl, bm_sn_fu, bm_sn_stats_bl, bm_sn_stats_fu
)

# Duplicate the GAMMs to get seperate p-values
gamm_memory_lft <- gamm_compare_em_memory %>%
  select(p_value_lrt)

gamm_memory_gamm <- gamm_compare_em_memory %>%
  select(p_value_gamm)

gamm_hc_lft <- gamm_compare_em_adj_hc %>%
  select(p_value_lrt)

gamm_hc_gamm <- gamm_compare_em_adj_hc %>%
  select(p_value_gamm)

# Model Comparisons and General Analyses
model_comp_general_models <- list(
  gamm_memory_lft, gamm_memory_gamm,
  gamm_hc_lft, gamm_hc_gamm,
  hc_age_lme, mem_age_lme
)

# Define a function to extract p-values
extract_p_value <- function(model) {
  if ("p_value" %in% names(model)) {
    return(model[["p_value"]])
  } else if ("p_value_gamm" %in% names(model)) {
    return(model[["p_value_gamm"]])
  } else if ("p_value_lrt" %in% names(model)) {
    return(model[["p_value_lrt"]])
  } else {
    return(NULL)
  }
}

adjust_p_values <- function(models, model_names) {
  p_values <- unlist(lapply(models, extract_p_value))
  p_adjusted <- p.adjust(p_values, method = "fdr")

  data.frame(
    model_name = model_names,
    original_p_value = format(p_values, digits = 3, scientific = FALSE, na.encode = TRUE),
    adjusted_p_value = format(p_adjusted, digits = 3, scientific = FALSE, na.encode = TRUE)
  )
}

# Adjust p-values for each group
memory_hippocampus_p_values <- adjust_p_values(memory_hippocampus_edu_models, c("emhp_merged", "model_baseline", "model_followup", "edu_baseline", "edu_followup", "edu_long"))
education_p_values <- adjust_p_values(education_models, c("edu_cr", "edu_bm", "edu_cr_int", "edu_bm_int"))
cr_bm_p_values <- adjust_p_values(cr_bm_models, c("cr_validation", "bm_age_lme", "cr_age_lme"))
functional_connectivity_long_p_values <- adjust_p_values(functional_connectivity_long, c("bm_dmn_stats", "bm_dmn", "bm_ecn_stats", "bm_ec", "bm_sn_stats", "bm_sn", "cr_dmn_stats", "cr_dmn", "cr_ecn_stats", "cr_ec", "cr_sn_stats", "cr_sn"))
functional_connectivity_cs_p_values <- adjust_p_values(functional_connectivity_cs, c("cr_dmn_bl", "cr_dmn_fu", "cr_dmn_stats_bl", "cr_dmn_stats_fu", "cr_ec_bl", "cr_ec_fu", "cr_ecn_stats_bl", "cr_ecn_stats_fu", "cr_sn_bl", "cr_sn_fu", "cr_sn_stats_bl", "cr_sn_stats_fu", "bm_dmn_bl", "bm_dmn_fu", "bm_dmn_stats_bl", "bm_dmn_stats_fu", "bm_ec_bl", "bm_ec_fu", "bm_ecn_stats_bl", "bm_ecn_stats_fu", "bm_sn_bl", "bm_sn_fu", "bm_sn_stats_bl", "bm_sn_stats_fu"))
model_comp_general_p_values <- adjust_p_values(model_comp_general_models, c("gamm_memory_lft", "gamm_memory_gamm", "gamm_hc_lft", "gamm_hc_gamm", "hc_age_lme", "mem_age_lme"))

# Combine all results
p_value_df <- rbind(
  memory_hippocampus_p_values,
  education_p_values,
  cr_bm_p_values,
  functional_connectivity_long_p_values,
  functional_connectivity_cs_p_values,
  model_comp_general_p_values
)

# Output the combined p-values
write.csv(p_value_df, file = "/tsd/p274/data/durable/projects/p027-cr_bm/p_value.csv")
```



```{r}
library(ggnewscale)

clean_data <- clean_data %>%
  mutate(
    dist = ifelse(dist_to_bm_line < dist_to_cr_line, "BM", "CR"),
    dist_cr_or_bm = ifelse(dist_to_bm_line < dist_to_cr_line, reverse_bm, -reverse_cr)
  )

bm_data <- clean_data %>% 
  filter(dist == "BM")

cr_data <- clean_data %>% 
  filter(dist == "CR")

test <- ggplot() +
  # Plot the first dataset
  geom_point(data = bm_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled, color = reverse_bm), size = 2) +
  scale_color_gradientn(
    colors = custom_colors_bm,
    limits = c(0.55, 2.59),
    values = c(0, 1/3, 2/3, 1),
    na.value = NA,
    name = "Brain Maintenance"
  ) +
  geom_smooth(data = bm_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled), method = "lm", color = "grey", fill = "grey") +
  
  # Add new color scale
  new_scale_color() +
  
  # Plot the second dataset
  geom_point(data = cr_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled, color = reverse_cr), size = 2) +
  scale_color_gradientn(
    colors = custom_colors_cr,
    limits = c(0.55, 2.59),
    values = c(0, 1/3, 2/3, 1),
    na.value = NA,
    name = "Cognitive Reserve"
  ) +
  geom_smooth(data = cr_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled), method = "lm", color = "grey", fill = "grey") +
  labs(
    y = "Memory Annual Change",
    x = "Hippocampus Annual Change"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  xlim(-1, 1) +
  ylim(-1, 1)

bm_data <- clean_data %>% 
  filter(closer_to == "BM")

cr_data <- clean_data %>% 
  filter(closer_to == "CR")

test2 <- ggplot() +
  # Plot the first dataset
  geom_point(data = bm_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled, color = reverse_bm), size = 2) +
  scale_color_gradientn(
    colors = custom_colors_bm,
    limits = c(0.55, 2.59),
    values = c(0, 1/3, 2/3, 1),
    na.value = NA,
    name = "Brain Maintenance"
  ) +
  geom_smooth(data = bm_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled), method = "lm", color = "grey", fill = "grey") +
  
  # Add new color scale
  new_scale_color() +
  
  # Plot the second dataset
  geom_point(data = cr_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled, color = reverse_cr), size = 2) +
  scale_color_gradientn(
    colors = custom_colors_cr,
    limits = c(0.55, 2.59),
    values = c(0, 1/3, 2/3, 1),
    na.value = NA,
    name = "Cognitive Reserve"
  ) +
  geom_smooth(data = cr_data, aes(x = res_hc_slopes_scaled, y = res_mem_slopes_scaled), method = "lm", color = "grey", fill = "grey") +
  labs(
    y = "Memory Annual Change",
    x = "Hippocampus Annual Change"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  xlim(-1, 1) +
  ylim(-1, 1)
```

```{r figx, echo=FALSE, dpi = 400, fig.width=6, fig.height=4, message=FALSE, warning=FALSE}
test
test2
```
